### 2019-01-09 archive/accessPEP_summary.Rmd
archiving older files

---
title: "Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar"
author: ""
date: 
output:
  html_document: default
  pdf_document:
    includes:
      in_header: mystyle.sty
  word_document:
    reference_docx: format_docs/word_styles_reference_01.docx
css: style.css
---

## Goals
- To estimate burden of rabies in Madagascar using nationally available data from clinics that provide PEP.
- To explore the potential impacts of expanded access to PEP

## Data
- Patient data from 19 ARMC accross the country over four years (2013 - 2017)
  - includes dates and home locations (to the district level) of patient reporting for PEP
- Patient data from Moramanga ARMC over 28 months (Sep 2016 - Dec 2018)
  - includes dates and home locations (to the commune level) of patients reporting for PEP
  - also includes rabies exposure status based on a combination of clinic-based triage and contact
  tracing (Rajeev et al. 2018)
- Travel times estimated from the Malaria Atlas Project's global friction surface for 2015
  - We use the global friction surface for 2015 generated by the Malaria Atlas Project ( https://map.ox.ac.uk/research-project/accessibility_to_cities/, Weiss et al. 2015,) and GPS points of clinics to get travel time to the nearest ARMC for the country of Madagascar at a 1 x 1 km scale. 
  - We then calculated a weighted average by human population to the commune or district level, using administrative shapefiles available trhough the UN Office for the Coordination of Humanitarian Affairs. 
  - For each clinic, we defined the catchment area as all communes for which the clinic was the closest ARMC.
- Human population estimates were taken from the 2015 UN adjusted population projections from World Pop (www.worldpop.org, Linaird et al. 2012) and also aggreagated to the designated admin level.

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
rm(list = ls())
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
select <- dplyr::select
knitr::opts_chunk$set(fig.pos = 'h', echo = FALSE, message = FALSE, warning = FALSE, results = 'hide')
```

```{r CTAR data}

```

## Cleaning the data
```{r get data}
## read in data
ctar_data <- read.csv("data/SaisieRage_DATA_2018-09-21_1755.csv")
ctar_metadata <- read.csv("data/ctar_metadata.csv")
mada_district <- readOGR("data/MadaGIS/MadaPops.shp")
dist_catchments <- read.csv("output/catchments_district_unmasked.csv", row.names = 1)
names(dist_catchments) <- c("mdg_dis_co", "CTAR", "ttimes_weighted")

## Centers with no data
no_data <- c("IPM", "Fianarantsoa", "Ambatomainty", "Ambovombe Androy",
             "Antsiranana", "Mandritsara", "Marolambo", "Nosy be", "Sainte Marie",
             "Taolagnaro", "Tsiroanomandidy", "Vangaindrano")
ctar_metadata$exclude <- 0
ctar_metadata$exclude[ctar_metadata$CTAR %in% no_data] <- 1

## Get color palette from other network fig!
img <- readJPEG("~/Downloads/Dp0j4GNUwAAPPJV.jpeg")
pal <- image_palette(img, 31)
add.alpha <- function(col, alpha=1){
  if(missing(col))
    stop("Please provide a vector of colours.")
  apply(sapply(col, col2rgb)/255, 2, 
                     function(x) 
                       rgb(x[1], x[2], x[3], alpha=alpha))  
}
ctar_metadata$color <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                 "#63FFAC", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                 "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                 "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                 "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2")
ctar_metadata$fill <- ctar_metadata$color
ctar_metadata$fill[ctar_metadata$CTAR %in% no_data] <- "#D3D3D3"
ctar_metadata$fill <- add.alpha(ctar_metadata$fill, 0.25)
## Getting ctar district and commune
ctar_metadata$ctar_dist <- mada_district$mdg_dis_co[match(ctar_metadata$District, 
                                                           mada_district$district)]
pts <- SpatialPoints(cbind(ctar_metadata$LONGITUDE, ctar_metadata$LATITUDE), 
                   proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
ctar_metadata$ctar_comm <- over(pts, mada_communes)$mdg_cm_

display_palette(ctar_metadata$color)
dist_catchments %>% left_join(select(ctar_metadata, CTAR, exclude, 
                                     color, fill, ctar_dist, LATITUDE, LONGITUDE)) -> dist_catchments
mada_district@data <- left_join(mada_district@data, dist_catchments)
dist_catchments$ctar_in <- ifelse(dist_catchments$mdg_dis_co %in% ctar_metadata$ctar_dist, 1, 0)
```
1. Dealing with underreporting at the clinic level
2. Matching names to the commune level
3. Estimating catchments and weighted travel times by admin level
```{r catchment figure}
## get exposure matrix
ctar_data %>% 
  group_by(district, id_ctar) %>% # group by district and ctar
  summarize(n = n()) -> exp_mat # get count of exposures
dist_points <- cbind(select(mada_district@data,
                            mdg_dis_co), coordinates(mada_district))
names(dist_points)[2:3] <- c("dist_long", "dist_lat")
dist_points %>%
 left_join(select(ctar_metadata, ctar_dist, LATITUDE, LONGITUDE), 
           by = c("mdg_dis_co" = "ctar_dist")) %>%
 mutate(dist_long = coalesce(dist_long, LONGITUDE), 
        dist_lat = coalesce(dist_lat, LATITUDE)) -> dist_points

exp_mat %>% 
  left_join(select(dist_points, mdg_dis_co, dist_long, dist_lat), 
            by = c("district" = "mdg_dis_co")) %>%
  left_join(select(ctar_metadata, id_ctar, LATITUDE, LONGITUDE,
                   color, fill, ctar_dist)) -> dist_lines
dist_lines$size <- log(dist_lines$n + 0.1)*0.2

ctar_data %>%
  group_by(district) %>%
  summarize(n = n()) %>%
  right_join(select(dist_points, mdg_dis_co, dist_long, dist_lat),
             by = c("district" = "mdg_dis_co")) %>%
  left_join(dist_catchments, by = c("district" = "mdg_dis_co") ) -> dist_points
dist_points$size <- log(dist_points$n + 0.1)*0.25

pdf("figs/network_1.pdf")
plot(mada_district, col = mada_district$fill, border = "grey") 
points(jitter(dist_lines$dist_long, factor = 2), jitter(dist_lines$dist_lat, factor = 2)
       , pch = 1, cex = dist_points$size, 
       col = dist_lines$color)
segments(dist_lines$LONGITUDE, dist_lines$LATITUDE, x1 = dist_lines$dist_long, y1 = dist_lines$dist_lat,
         lwd = dist_lines$size*2, col = dist_lines$color)
dev.off()

legend("topleft", c(">100", ">200", ">400", ">800"), title = "Number of total reported bites", 
       text.width = 3, pt.cex = circle_size, col = "grey",  pch = 16, inset = c(0.1, 0.1), lwd = 1.5, bty = "n", xpd = TRUE, y.intersp=2)

exp_mat %>% 
  left_join(select(ctar_metadata, CTAR, id_ctar)) %>%
  left_join(select(dist_catchments, mdg_dis_co, CTAR, exclude), by = c("district" = "mdg_dis_co")) %>%
  group_by(district) %>%
  mutate(prop = n/sum(n),
         catch_match = ifelse(CTAR.x == CTAR.y, 1, 0)) -> catch_check
hist(catch_check$prop)
ggplot(data = catch_check, aes(CTAR.x, district)) + 
  geom_tile(aes(fill = prop)) +
  scale_fill_continuous(type = "viridis", name = "proportion") +
  geom_text(label = catch_check$catch_match)

```

```{r}
## get vials delivered
ctar_metadata %>% 
  dplyr::select(id_ctar, starts_with("doses")) %>%
  gather(year, vials_delivered, -id_ctar) %>%
  mutate(year = as.numeric(substr(as.character(year), 7, 10))) -> vials_delivered

## getting dose time series by clinic
########### Need to fix dates first! ###############
ctar_data %>%
  select(id_ctar, datej0, datej3, datej7, datej28) %>%
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>%
  gather(dose, date, -id_ctar) %>%
  group_by(id_ctar, date) %>%
  summarise(n = n()) -> dose_ts
dose_ts <- dose_ts[!is.na(dose_ts$date), ]
dose_ts$ctar <- ctar_metadata$CTAR[match(dose_ts$id_ctar, ctar_metadata$id_ctar)]
dose_ts <- dose_ts[!is.na(dose_ts$ctar), ]

## getting theoretical # vials by year
dose_ts %>%
  mutate(year = year(date), vials = ceiling(n/2)) %>%
  group_by(id_ctar, year) %>%
  summarize(vials_estimated = sum(vials)) %>%
  left_join(vials_delivered) -> dose_comp

######### Patient time series
ctar_data %>%
  select(id_ctar, date_de_consultation) %>%
  mutate(date_de_consultation = ymd(date_de_consultation)) %>%
  gather(dose, date_de_consultation, -id_ctar) %>%
  group_by(id_ctar, date_de_consultation) %>%
  summarise(n = n()) -> patient_ts
patient_ts <- patient_ts[!is.na(patient_ts$date_de_consultation), ]
patient_ts$ctar <- ctar_metadata$CTAR[match(patient_ts$id_ctar, ctar_metadata$id_ctar)]
patient_ts <- patient_ts[!is.na(patient_ts$ctar), ]

ggplot(data = patient_ts, aes(date_de_consultation, ctar)) + 
  geom_tile(aes(fill = n)) +
  scale_fill_continuous(type = "viridis", name = "Patients")

patient_ts %>%
  filter(id_ctar == 1) %$%
  hist(n, breaks = 100, col = "blue", 
       xlab = "Patients reporting", main = "Soanierana Ivongo")
patient_ts %>%
  filter(ctar == "Toamasina") %$%
  hist(n, breaks = 100, 
       xlab = "Patients reporting", main = "Toamasina", col = "red")

## getting reporting
start_date <- ymd("2014-01-01")
end_date <- ymd("2017-12-31")
ts <- as_tibble(seq(start_date, end_date, by = "day"))

patient_ts %>%
  right_join(ts, by = c("date_de_consultation" = "value")) %>%
  select(id_ctar, date_de_consultation, n) %>%
  spread(id_ctar, n) %>%
  replace(., is.na(.), 0) -> doses_wide
dose_mat <- as.matrix(doses_wide[, 2:(ncol(doses_wide) - 1)])

date_mat <- matrix(NA, nrow(dose_mat), ncol(dose_mat))
for (j in 1:ncol(date_mat)){
  rle(dose_mat[ , j]) %>%
    unclass() %>%
    as.data.frame() %>%
    mutate(end = cumsum(lengths),
         start = c(1, lag(end)[-1] + 1)) %>%
    filter(values == 0, lengths >= 10) -> rles
  for (i in 1:nrow(rles)){
    date_mat[rles$start[i]:rles$end[i], j] <- 0
  }
}

date_mat <- replace(date_mat, is.na(date_mat), 1)
date_mat %>% 
  as_tibble() %>%
  group_by(date = year(doses_wide$date_de_consultation)) %>%
  summarise_all(funs(sum(.)/n())) -> clinic_reporting
colnames(clinic_reporting) <- colnames(doses_wide)[1:(ncol(doses_wide)-1)]
clinic_reporting <- gather(clinic_reporting, id_ctar, prop, -date_de_consultation)
clinic_reporting$ctar <- ctar_metadata$CTAR[match(clinic_reporting$id_ctar, 
                                                  ctar_metadata$id_ctar)]
clinic_reporting$id_ctar <- as.numeric(clinic_reporting$id_ctar)

ggplot(clinic_reporting, aes(x = date_de_consultation, y = ctar, fill = prop)) + 
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue")
```
4. Contacts + other issues (i.e. date errors)
5. Corrected incidence

## Models of bites
We assume that the incidence of bites at the district/commune level, reflects differences in access to PEP. 
We fit a generalized linear model with the annual average bites modeled as a rate per population (with a log offset oft the population) for three datasets:
1. National data corrected and averaged to the district level
2. National data corrected and averaged to the commune level
3. Moramanga data

### Results of models
#### Fit
#### Compare relationships

## Estimating burden
We used our model to predict average annual bite incidence for all 114 districts in Madagascar, and estimated deaths due to rabies.

Using the standard decision tree framework, we assumed that reported bites were proportional to rabies incidence, and calculated deaths, assuming uniform reporting across Madagascar (taken from Moramanga paper = 0.85):
$$
  deaths_{i} = (frac{{p_{rabid} \times B_{i}}{\rho} - p_{rabid} \times B_{i}) \times p_{death}
$$

Alternaetively, we calculated deaths based on assumptions about endemic incidence of rabies exposures:
$$
  deaths_{i} = (R_{i} - ({p_{rabid} \times B_{i})) \times p_{death}
$$

where $R_{i}$ is drawn from a uniform distribution between the minimum and maximum expected number of human exposures. We calculate this as:

<br>

$$r \times D \times p_{exp}$$

<br>

where $r$ is rabies incidence in dogs in the absence of any vaccination multiplied by the estimated dog population in the commune ($D_i$) and the exposure rate per rabid dogs ($p_{exp}$ = 0.39 persons exposed per rabid dog)(Hampson et al. 2018). We estimate the dog population by using a human:dog ratio (HDR) of 5 to generate our maximum expected incidence and an HDR of 25 for our minimum expected incidence. As there is little data on dog populations in Madagascar, this range of HDRs encompasses a wide range observed across Africa(cite?).

The proportion of reported bites that are rabies exposures ($p_{rabid}$) are also constrained by the number of rabies exposures:
$$
           p_{rabid}=
            \begin{cases}
            x, & \text{if}\ \frac{R_i \times \rho_{max}}{B_i} > x \\
            \frac{R_i}{B_i}, & \text{otherwise}
            \end{cases}
$$
<br>

such that rabid reported bites (i.e. $p_{rabid}\timesB_{i}$) cannot exceed the expected number of human exposures given maximum reporting (i.e. $R_{i} \times \rho_{max}$. The range of $p_{rabid}$ and $\rho_max$ are taken from Rajeev et al. 2018 on contact tracing data from the Moramanga District.

Finally, we also accounted for differences in reporting probability of rabid vs. non-rabid bites as a function of travel times. Using the Moramanga data, we looked at how travel times predict the $p_{rabid}$ so that:
model here

In our estimation of burden, we predict what p_rabid is but also similarly constrain it so that rabid reported bites cannot exceed total reported bites.

## Estimating the impact of expanding PEP Access
We use this framework to compare three scenarios of PEP provisioning in Madagascar:
1. The baseline with the current clinic locations (n = 31)
2. Expansion of ARMC to one clinic per district (n = 114)
3. Expansion of ARMC to all CSB IIs (n = ...?)

We use data on the location of **CSBs** provided by IPM to regenerate travel times to the nearest ARMC given expansion as per scenario 2 and 3. We then predict the expected bite incidence from the model given these new travel times and compare the relative decreases in burden for the three scenarios.

<br>

### Sensitivity Analyses
- Sensitivity to assumptions of R
  - pop?
- Sensitivity to assumptions of p_rabid and p_max

```{r catchment figure}
## get exposure matrix
ctar_data %>% 
  group_by(district, id_ctar) %>% # group by district and ctar
  summarize(n = n()) -> exp_mat # get count of exposures
dist_points <- cbind(select(mada_district@data,
                            mdg_dis_co), coordinates(mada_district))
names(dist_points)[2:3] <- c("dist_long", "dist_lat")
dist_points %>%
 left_join(select(ctar_metadata, ctar_dist, LATITUDE, LONGITUDE), 
           by = c("mdg_dis_co" = "ctar_dist")) %>%
 mutate(dist_long = coalesce(dist_long, LONGITUDE), 
        dist_lat = coalesce(dist_lat, LATITUDE)) -> dist_points

exp_mat %>% 
  left_join(select(dist_points, mdg_dis_co, dist_long, dist_lat), 
            by = c("district" = "mdg_dis_co")) %>%
  left_join(select(ctar_metadata, id_ctar, LATITUDE, LONGITUDE,
                   color, fill, ctar_dist)) -> dist_lines
dist_lines$size <- log(dist_lines$n + 0.1)*0.2

ctar_data %>%
  group_by(district) %>%
  summarize(n = n()) %>%
  right_join(select(dist_points, mdg_dis_co, dist_long, dist_lat),
             by = c("district" = "mdg_dis_co")) %>%
  left_join(dist_catchments, by = c("district" = "mdg_dis_co") ) -> dist_points
dist_points$size <- log(dist_points$n + 0.1)*0.25

pdf("figs/network_1.pdf")
plot(mada_district, col = mada_district$fill, border = "grey") 
points(jitter(dist_lines$dist_long, factor = 2), jitter(dist_lines$dist_lat, factor = 2)
       , pch = 1, cex = dist_points$size, 
       col = dist_lines$color)
segments(dist_lines$LONGITUDE, dist_lines$LATITUDE, x1 = dist_lines$dist_long, y1 = dist_lines$dist_lat,
         lwd = dist_lines$size*2, col = dist_lines$color)
dev.off()

legend("topleft", c(">100", ">200", ">400", ">800"), title = "Number of total reported bites", 
       text.width = 3, pt.cex = circle_size, col = "grey",  pch = 16, inset = c(0.1, 0.1), lwd = 1.5, bty = "n", xpd = TRUE, y.intersp=2)

exp_mat %>% 
  left_join(select(ctar_metadata, CTAR, id_ctar)) %>%
  left_join(select(dist_catchments, mdg_dis_co, CTAR, exclude), by = c("district" = "mdg_dis_co")) %>%
  group_by(district) %>%
  mutate(prop = n/sum(n),
         catch_match = ifelse(CTAR.x == CTAR.y, 1, 0)) -> catch_check
hist(catch_check$prop)
ggplot(data = catch_check, aes(CTAR.x, district)) + 
  geom_tile(aes(fill = prop)) +
  scale_fill_continuous(type = "viridis", name = "proportion") +
  geom_text(label = catch_check$catch_match)

```

```{r}
## get vials delivered
ctar_metadata %>% 
  dplyr::select(id_ctar, starts_with("doses")) %>%
  gather(year, vials_delivered, -id_ctar) %>%
  mutate(year = as.numeric(substr(as.character(year), 7, 10))) -> vials_delivered

## getting dose time series by clinic
########### Need to fix dates first! ###############
ctar_data %>%
  select(id_ctar, datej0, datej3, datej7, datej28) %>%
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>%
  gather(dose, date, -id_ctar) %>%
  group_by(id_ctar, date) %>%
  summarise(n = n()) -> dose_ts
dose_ts <- dose_ts[!is.na(dose_ts$date), ]
dose_ts$ctar <- ctar_metadata$CTAR[match(dose_ts$id_ctar, ctar_metadata$id_ctar)]
dose_ts <- dose_ts[!is.na(dose_ts$ctar), ]

## getting theoretical # vials by year
dose_ts %>%
  mutate(year = year(date), vials = ceiling(n/2)) %>%
  group_by(id_ctar, year) %>%
  summarize(vials_estimated = sum(vials)) %>%
  left_join(vials_delivered) -> dose_comp

######### Patient time series
ctar_data %>%
  select(id_ctar, date_de_consultation) %>%
  mutate(date_de_consultation = ymd(date_de_consultation)) %>%
  gather(dose, date_de_consultation, -id_ctar) %>%
  group_by(id_ctar, date_de_consultation) %>%
  summarise(n = n()) -> patient_ts
patient_ts <- patient_ts[!is.na(patient_ts$date_de_consultation), ]
patient_ts$ctar <- ctar_metadata$CTAR[match(patient_ts$id_ctar, ctar_metadata$id_ctar)]
patient_ts <- patient_ts[!is.na(patient_ts$ctar), ]

ggplot(data = patient_ts, aes(date_de_consultation, ctar)) + 
  geom_tile(aes(fill = n)) +
  scale_fill_continuous(type = "viridis", name = "Patients")

patient_ts %>%
  filter(id_ctar == 1) %$%
  hist(n, breaks = 100, col = "blue", 
       xlab = "Patients reporting", main = "Soanierana Ivongo")
patient_ts %>%
  filter(ctar == "Toamasina") %$%
  hist(n, breaks = 100, 
       xlab = "Patients reporting", main = "Toamasina", col = "red")

## getting reporting
start_date <- ymd("2014-01-01")
end_date <- ymd("2017-12-31")
ts <- as_tibble(seq(start_date, end_date, by = "day"))

patient_ts %>%
  right_join(ts, by = c("date_de_consultation" = "value")) %>%
  select(id_ctar, date_de_consultation, n) %>%
  spread(id_ctar, n) %>%
  replace(., is.na(.), 0) -> doses_wide
dose_mat <- as.matrix(doses_wide[, 2:(ncol(doses_wide) - 1)])

date_mat <- matrix(NA, nrow(dose_mat), ncol(dose_mat))
for (j in 1:ncol(date_mat)){
  rle(dose_mat[ , j]) %>%
    unclass() %>%
    as.data.frame() %>%
    mutate(end = cumsum(lengths),
         start = c(1, lag(end)[-1] + 1)) %>%
    filter(values == 0, lengths >= 10) -> rles
  for (i in 1:nrow(rles)){
    date_mat[rles$start[i]:rles$end[i], j] <- 0
  }
}

date_mat <- replace(date_mat, is.na(date_mat), 1)
date_mat %>% 
  as_tibble() %>%
  group_by(date = year(doses_wide$date_de_consultation)) %>%
  summarise_all(funs(sum(.)/n())) -> clinic_reporting
colnames(clinic_reporting) <- colnames(doses_wide)[1:(ncol(doses_wide)-1)]
clinic_reporting <- gather(clinic_reporting, id_ctar, prop, -date_de_consultation)
clinic_reporting$ctar <- ctar_metadata$CTAR[match(clinic_reporting$id_ctar, 
                                                  ctar_metadata$id_ctar)]
clinic_reporting$id_ctar <- as.numeric(clinic_reporting$id_ctar)

ggplot(clinic_reporting, aes(x = date_de_consultation, y = ctar, fill = prop)) + 
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue")
```

## Submission of forms
- Missingness seems to have been due to periods of time without form submission
- Calculating periods of time

```{r clean data}
## districts that are non-reporter and data not available yet CTAR
mada_district@data %>%
select(district = mdg_dis_co, pop = pop2015adj, ttimes_weighted, CTAR, exclude) -> covars

to_exclude <- c(grep("contact", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE),
                grep("contam", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE),
                grep("consom", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE),
                grep("passage", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE))
to_exclude <- to_exclude[-grep("date", to_exclude, ignore.case = TRUE)]
ctar_data$excl <- 0
ctar_data$excl[ctar_data$remarque %in% to_exclude] <- 1

ctar_data %>% 
  filter(excl != 1) %>% # exclude contacts
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>% # format dates
  group_by(year = year(date_de_consultation), district, id_ctar) %>% # group by year and district
  summarize(n = n()) -> exposures # get count of exposures

exposures %>%
  rename(date = year) %>%
  left_join(covars) %>% # add pop + ttimes + reporting
  left_join(clinic_reporting) %>%   # add reporting by clinic + year
  filter(exclude == 0, prop > 0.25) %>%
  group_by(district, date) %>%
  mutate(bites = n/prop) %>%
  summarise(bites = sum(bites, na.rm = TRUE)) -> exps_cleaned 
exps_cleaned %>%
  group_by(district) %>%
  summarise(bites = mean(bites)) %>%
  left_join(covars) -> exps_cleaned

library(lme4)

## poisson with total pop as offset
mod <- glm(round(bites) ~ ttimes_weighted + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)
par(mfrow = c(2, 1), mar = c(4, 4, 1, 1))
plot(exps_cleaned$bites, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Observed bites", ylab = "Predicted bites")
abline(a = 0, b = 1, lty = 2, col = "blue")
legend("topright", col = "blue", legend = "1:1 line", lty = 2, bty = "n", seg.len = 1, 
       inset = c(0.0, -0.1))
plot(exps_cleaned$ttimes_weighted/60, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Travel times (hrs)", ylab = "Bites", col = alpha("blue", 0.7), 
     ylim = c(0, max(exps_cleaned$bites)))
points(exps_cleaned$ttimes_weighted/60, exps_cleaned$bites, bty = "l", pch = 20,
     xlab = "Travel times (hrs)", col = alpha("darkred", 0.7))
legend("topright", col = c(alpha("blue", 0.7), alpha("darkred", 0.7)), legend = c("Predicted", "Observed"), 
       pch = 20, bty = "n")
```

```{r functions for getting burden}
get.burden <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho_max = 0.9,
                       max_HDR = 25, min_HDR = 5, 
                       dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- store_p_rabid <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    mean_exps <- runif(length(pop), min = human_exp_rate*dog_rabies_inc*pop/max_HDR, 
                    max = human_exp_rate*dog_rabies_inc*pop/min_HDR)
    rabid_exps <- rpois(length(mean_bites), mean_exps)
    p_rabid_t <- ifelse(bites*p_rabid/rabid_exps > rho_max, (rho_max*rabid_exps)/bites, p_rabid)
    deaths <- rbinom(length(mean_bites), rabid_exps - round(bites*p_rabid_t), prob_death)
    store_deaths[, i] <- deaths
    store_p_rabid[, i] <- p_rabid_t
  }

  results <- cbind(get.meanCI(store_deaths), get.meanCI(store_p_rabid))
  colnames(results) <- c("deaths_mean", "deaths_lowerCI", "deaths_upperCI",
                         "p_rabid_mean", "p_rabid_lowerCI", "p_rabid_upperCI")
  results <- cbind(covars, results)
  return(results)
}

get.burden.flat <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho = 0.85, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    unreported <- bites*p_rabid/rho - bites*p_rabid
    deaths <- rbinom(length(mean_bites), round(unreported), prob_death)
    store_deaths[, i] <- deaths
  }
  results <- get.meanCI(store_deaths)
  colnames(results) <- c("deaths_mean", "death_lowerCI", "deaths_upperCI")
  results <- cbind(covars, results)
  return(results)
}

get.meanCI <- function(matrix_by_row, nboots = 1000) {
    mean_rows <- apply(matrix_by_row, 1, mean)
    CIs <- apply(matrix_by_row, 1, boots, nboot = nboots)
    lower <- unlist(CIs)[1, ] 
    upper <- unlist(CIs)[2, ]
    summ <- cbind(mean_rows, lower, upper)
    colnames(summ) <- c("mean", "lower", "upper")
    return(summ)
}

boots <- function(x, nboot = 1000){
  n = length(x)
  xbar = mean(x)
  tmpdata = sample(x,n*nboot, replace=TRUE)
  bootstrapsample = matrix(tmpdata, nrow=n, ncol=nboot)
  bsmeans = colMeans(bootstrapsample)
  deltastar = bsmeans - xbar
  d = quantile(deltastar, c(0.05, 0.95))
  ## Get from quantiles
  ci = xbar - c(d[2], d[1])

  ## Sorted CIs
  sorteddeltastar = sort(deltastar)
  # # Look at the sorted results
  # Find the .05 and .95 critical values of deltastar
  d9alt = sorteddeltastar[5]
  d1alt = sorteddeltastar[950]
  # Find and print the 80% confidence interval for the mean
  ciAlt = xbar - c(d1alt,d9alt)
  # cat(’Alternative confidence interval: ’,ciAlt, ’\n’)
  return(ci)
}

```

```{r}
results_pr20 <- get.burden(p_rabid = 0.20, rho_max = 0.98)
results_pr60 <- get.burden(p_rabid = 0.60,rho_max = 0.98)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 20)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_base <- get.burden(p_rabid = 0.40, rho_max = 0.98)

p_rabid_check <- seq(0.10, 0.80, by = 0.1)
check <- data.frame(p_rabid = 0, district = NA, ttimes = 0, deaths_mean = 0)
for (i in 1:length(p_rabid_check)) {
  print(paste(i, "/", length(p_rabid_check)))
  results <- get.burden(p_rabid = p_rabid_check[i], rho_max = 0.98)
  p_rabid <- rep(p_rabid_check[i], length(results$deaths_mean))
  check_one <- cbind(p_rabid, results$district, results$ttimes, results$deaths_mean)
  colnames(check_one) <- c("p_rabid", "district", "ttimes", "deaths_mean")
  check <- rbind(check, check_one)
}

check <- check[-1, ]
check$base <- rep(results_base$deaths_mean, length(p_rabid_check))
ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = deaths_mean/base)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 1, name = "Mean deaths")

rho_max_check <- seq(0.90, 0.99, by = 0.01)
check <- data.frame(rho_max = 0, district = NA, ttimes = 0, deaths_mean = 0)
for (i in 1:length(rho_max_check)) {
  print(paste(i, "/", length(rho_max_check)))
  results <- get.burden(p_rabid = 0.4, rho_max = rho_max_check[i])
  rho_max <- rep(rho_max_check[i], length(results$deaths_mean))
  check_one <- cbind(rho_max, results$district, results$ttimes, results$deaths_mean)
  colnames(check_one) <- c("rho_max", "district", "ttimes", "deaths_mean")
  check <- rbind(check, check_one)
}

check <- check[-1, ]
check$base <- rep(results_base$deaths_mean, length(rho_max_check))
check %>%
  group_by(district) %>%
  mutate(min = min(deaths_mean), max = max(deaths_mean), 
         standardized = (deaths_mean - base)/(max - base)) -> check

ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = standardized)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 0, name = "Mean deaths") 
ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = deaths_mean)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 0, name = "Mean deaths")

mada_dist_gg <- tidy(mada_district, region = "mdg_dis_co")
mada_dist_gg <- left_join(mada_dist_gg, exps_cleaned, by = c("id" = "district"))

ggplot() +
  geom_polygon(data = mada_dist_gg, aes(x = long, y = lat, group = group, fill = CTAR)) +
  scale_colour_viridis_d()

```

## Commune level mods
```{r commune matching, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
commune_matching <- read.csv("data/commune_check_CH.csv")
mada_communes <- readOGR("output/communes/communes_extracted.shp")
mada_communes$match <- tolower(str_replace_all(mada_communes$commune,"[^[:graph:]]", "/"))

mada_communes@data %>%
  select(match, mtch_ds, mdg_cm_) %>%
  left_join(commune_matching, .) %>%
  select(name, match, district, mdg_cm_) -> comm_matches

ctar_data$name <- tolower(str_replace_all(ctar_data$commune,"[^[:graph:]]", "/"))
left_join(ctar_data, comm_matches) %>%
  distinct(record_id, .keep_all = TRUE) -> ctar_data

## Looking at proportion matched
ctar_data %>%
  group_by(id_ctar) %>%
  summarise(n = n(), unmatched = sum(is.na(mdg_cm_)),
            prop_matched = unmatched/n) -> ctar_data_matched

ctar_metadata <- merge(ctar_metadata, ctar_data_matched)
ctar_metadata <- ctar_metadata[!is.na(ctar_metadata$id_ctar), ]
ggplot(ctar_metadata, aes(CTAR, prop_matched)) +
  geom_point() +
  coord_flip()  +
  ylab(label = "Proportion of patients matched to commune")
```

```{r }
## Ctar data at commune level
ctar_data %>% 
  filter(excl != 1) %>% # exclude contacts
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>% # format dates
  group_by(year = year(date_de_consultation), commune = mdg_cm_, id_ctar) %>% # group by year and district
  summarize(n = n()) -> exposures # get count of exposures

comm_catchments <- read.csv("output/catchments_commune_unmasked.csv", row.names = 1)
names(comm_catchments) <- c("mdg_cm_", "CTAR", "ttimes_weighted")
comm_catchments %>% left_join(select(ctar_metadata, CTAR, exclude, 
                                     color, fill, ctar_dist, LATITUDE, LONGITUDE)) -> comm_catchments
mada_communes@data <- left_join(mada_communes@data, comm_catchments)

mada_communes@data %>%
select(commune = mdg_cm_, pop = MDG__201, ttimes_weighted, CTAR, exclude) -> covars
mada_communes$ctar_in <- ifelse(mada_communes$mdg_cm_ %in% ctar_metadata$ctar_comm, 1, 0)

## Catchments with no data
no_data <- c("IPM", "Fianarantsoa", "Ambatomainty", "Ambovombe Androy",
             "Antsiranana", "Mandritsara", "Marolambo", "Nosy be", "Sainte Marie",
             "Taolagnaro", "Tsiroanomandidy", "Vangaindrano")
covars$exclude <- 0
covars$exclude[covars$CTAR %in% no_data] <- 1

exposures %>%
  rename(date = year) %>%
  left_join(covars) %>% # add pop + ttimes + reporting
  left_join(clinic_reporting) %>%   # add reporting by clinic + year
  filter(exclude == 0, prop > 0.25) %>%
  group_by(commune, date) %>%
  mutate(bites = n/prop) %>%
  summarise(bites = sum(bites, na.rm = TRUE)) -> exps_cleaned 
exps_cleaned %>%
  group_by(commune) %>%
  summarise(bites = mean(bites)) %>%
  left_join(covars) -> exps_cleaned

library(lme4)

## poisson with total pop as offset
exps_cleaned <- filter(exps_cleaned, ttimes_weighted != "Inf")
mod <- glm(round(bites) ~ ttimes_weighted + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)

par(mfrow = c(2, 1), mar = c(4, 4, 1, 1))
plot(exps_cleaned$bites, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Observed bites", ylab = "Predicted bites")
abline(a = 0, b = 1, lty = 2, col = "blue")
legend("topright", col = "blue", legend = "1:1 line", lty = 2, bty = "n", seg.len = 1, 
       inset = c(0.0, -0.1))
plot(exps_cleaned$ttimes_weighted/60, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Travel times (hrs)", ylab = "Bites", col = alpha("blue", 0.7), 
     ylim = c(0, max(exps_cleaned$bites)))
points(exps_cleaned$ttimes_weighted/60, exps_cleaned$bites, bty = "l", pch = 20,
     xlab = "Travel times (hrs)", col = alpha("darkred", 0.7))
legend("topright", col = c(alpha("blue", 0.7), alpha("darkred", 0.7)), legend = c("Predicted", "Observed"), 
       pch = 20, bty = "n")

ctar_data %>% 
  filter(excl != 1) %>% # exclude contacts
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>% # format dates
  group_by(commune = mdg_cm_, id_ctar) %>% # group by year and district
  summarize(n = n()) ->  exp_mat # get count of exposures

comm_catchments$exclude <- 0
comm_catchments$exclude[comm_catchments$CTAR %in% no_data] <- 1

exp_mat %>% 
  left_join(select(ctar_metadata, CTAR, id_ctar)) %>%
  left_join(select(comm_catchments, mdg_cm_, CTAR, exclude), by = c("commune" = "mdg_cm_")) %>%
  group_by(commune) %>%
  mutate(prop = n/sum(n),
         catch_match = ifelse(CTAR.x == CTAR.y, 1, 0)) -> catch_check
hist(catch_check$prop[catch_check$exclude == 0])
hist(catch_check$prop[catch_check$exclude == 1])

## getting var for ctar in
pts <- SpatialPoints(cbind(ctar_metadata$LONGITUDE, ctar_metadata$LATITUDE), 
                   proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
over(pts, mada_communes)$mtch_ds
```