### 2019-10-28 manuscript/summary_Oct24th.Rmd
prelim scripts added

---
title: "summary_Oct23"
author: "Malavika Rajeev"
date: "10/23/2019"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)

##' Libraries
library(rgdal)
library(raster)
library(malariaAtlas)
library(gdistance)
library(doRNG)
library(foreach)
library(dplyr)
library(geosphere)
select <- dplyr::select

## calling in functions
source("R/functions/access_functions.R")

## data
source("R/02_bitedata/03_estimate_biteinc.R") # either source this or output bite data
mada_communes <- readOGR("data/processed/shapefiles/mada_communes.shp")
mada_districts <- readOGR("data/processed/shapefiles/mada_districts.shp")
pop1x1<- raster("data/processed/rasters/worldpop2015adj_mada_1x1km.tif")
ctar_metadata <- read.csv("data/raw/ctar_metadata.csv")
friction_mada_masked <- raster("data/processed/rasters/friction_mada_masked.tif")

```

## The garden of forking paths...
Currently stuck proliferating options.

### Covariate options {.tabset}
```{r maps + rasters, echo = FALSE}
## Baseline surfaces
point_mat <- select(ctar_metadata, Y_COORD = LONGITUDE, X_COORD = LATITUDE)
# ttimes <- get.access(friction = friction_mada_masked, shapefile = mada_districts, 
#                      coords = point_mat, trans_matrix_exists = TRUE, 
#                      filename_trans = "data/processed/rasters/trans_gc_masked.rds", metric = "ttimes")
# writeRaster(ttimes, "data/processed/rasters/ttimes_baseline.tif")
# distance <- get.access(friction = friction_mada_masked, shapefile = mada_districts, 
#                       coords = point_mat, trans_matrix_exists = TRUE, 
#                        filename_trans = "data/processed/rasters/trans_gc_masked.rds", metric = "distance")
# writeRaster(distance, "data/processed/rasters/distance_baseline.tif")

ttimes <- raster("data/processed/rasters/ttimes_baseline.tif")
distance <- raster("data/processed/rasters/distance_baseline.tif")

library(rasterVis)
library(patchwork)

gplot(ttimes) +
  geom_raster(aes(fill = value/60)) +
  scale_fill_gradient(low = "white", high = "red", na.value = "white", 
                      name = "Travel times \n (hours)") +
  geom_point(data = point_mat, aes(x = Y_COORD, y = X_COORD), color = "blue", alpha = 0.25) +
  theme_void() +
  coord_equal()

gplot(distance) +
  geom_raster(aes(fill = value)) +
  scale_fill_gradient(low = "white", high = "red", na.value = "white", 
                      name = "Distance \n (km)") +
  geom_point(data = point_mat, aes(x = Y_COORD, y = X_COORD), color = "blue", alpha = 0.25) +
  theme_void() +
  coord_equal()

mada_districts@data %>%
  select(distcode, ctch_dsct, dist_cent) %>%
  mutate(from_long = coordinates(mada_districts)[, 1], 
         from_lat = coordinates(mada_districts)[, 2]) -> centroids_district
ctar_points <- select(ctar_metadata, CTAR, to_long = LONGITUDE, to_lat = LATITUDE)
segments_plot <- left_join(centroids_district, ctar_points, by = c("ctch_dsct" = "CTAR"))

gg_districts <- fortify(mada_districts, region = "distcode")

catch_cols <- as.character(ctar_metadata$color)
names(catch_cols) <- ctar_metadata$CTAR
```

```{r fig.width = 5, fig.height = 8}
ggplot() +
  geom_polygon(data = gg_districts, aes(x = long, y = lat, group = group), color = "grey50", 
               fill = NA) +
  geom_segment(data = segments_plot, aes(x = to_long, xend = from_long, y = to_lat,
                                          yend = from_lat, color = ctch_dsct)) +
  geom_point(data = segments_plot, aes(x = to_long, y = to_lat, color = ctch_dsct), size = 2) +
  geom_point(data = segments_plot, aes(x = from_long, y = from_lat, color = ctch_dsct), size = 2, 
             alpha = 0.5) +
  scale_color_manual(values = catch_cols, guide = "none") +
  theme_void() +
  labs(title = "Distance by centroid (km)")

```

### Resulting data
```{r}
allbites <- bind_rows(bites_by_distcent, bites_by_distwtd, bites_by_ttimes)

ggplot(allbites, aes(x = covar, y = avg_bites/pop*1e5, 
                            color = catchment)) +
  geom_point() +
  theme(legend.position = "none") +
  scale_color_manual(values = catch_cols, guide = "none") + 
  ylab("Annual bites per 100k") +
  xlab("Districts ordered by \n travel times (high to low)") +
  facet_wrap(~covar_name, scales = "free")

```

### Modeling options

#### Districts with & without CTAR
```{r ctar_bump, echo = FALSE}
ggplot(allbites, aes(x = as.factor(ctar_in_district), y = avg_bites/pop*1e5)) +
  geom_violin() +
  facet_wrap(~covar_name)
```

#### A Random Effect by Catchment
Models with a random intercept by catchment. May help to capture variation between clinics 
not explained by travel times alone. And provide more realistic estimates of vials.

```{r echo = FALSE}
ggplot(allbites, aes(x = as.factor(catchment), y = avg_bites/pop*1e5)) +
  geom_boxplot() +
  facet_wrap(~covar_name) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45))
```

## Model Results
### Convergence fig
```{r convergence}
model_ests <- read.csv("output/bitemod_results.csv")
## convergence plots
model_ests %>%
  select(data_source, covar_name, data_source, scale, pop_predict, ctar_bump, 
           summed, intercept, val = psrf_upper) %>%
  mutate(type = "psrf_upper") -> psrf
model_ests %>%
  select(data_source, covar_name, data_source, scale, pop_predict, ctar_bump, 
         summed, intercept, val = mpsrf) %>%
  mutate(type = "mpsrf") -> mpsrf
convergence <- bind_rows(mpsrf, psrf)
ggplot(convergence, aes(x = type, y = val, color = interaction(data_source, intercept))) + 
  geom_boxplot() +
  ylim(c(0.90, 1.2)) +
  facet_grid(pop_predict ~ covar_name, scales = "free_x", drop = TRUE) +
  geom_hline(yintercept = 1, linetype = 2, color = "grey")

```

### DIC fig
```{r}
## Estimates
model_ests %>%
  group_by(data_source, covar_name, data_source, scale, pop_predict, ctar_bump, 
           summed, intercept) %>%
  summarize(dic = mean(dic)) -> dic_ests
dic_ests %>%
  group_by(data_source, covar_name) %>%
  arrange(dic, .by_group = TRUE) -> dic_ranks
ggplot(data = dic_ranks, aes(x = pop_predict, y = dic, color = interaction(scale, intercept),
                             shape = ctar_bump))+
  geom_point() +
  facet_grid(data_source ~ covar_name, scales = "free") +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))
```

### Estimates
```{r}
model_ests %>%
  select(params, Mean, covar_name, pop_predict, intercept, ctar_bump, summed, data_source,
         scale) %>%
  spread(key = params, value = Mean, fill = 0) -> model_means

## Estimates of access
ggplot(data = filter(model_means, pop_predict != "onlyPop"), 
       aes(x = interaction(scale, data_source), y = beta_access, 
           color = intercept, shape = ctar_bump)) +
  geom_point(alpha = 0.5, size = 4) +
  scale_color_manual(values = c("darkred", "#cc7722", "#004b49"), name = "Scale") +
  facet_grid(covar_name ~ pop_predict, scales = "free", drop = TRUE)  +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))

```

### Predictions to Data
```{r}
preds_grouped <- read.csv("output/preds/preds_grouped_all.csv")

ggplot(data = filter(preds_grouped, intercept == "fixed"), 
       aes(x = log(avg_bites + 0.1), y = log(bites_mean + 0.1), 
                                 color = interaction(scale, data_source), shape = ctar_bump)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  scale_color_manual(values = c("darkred", "#cc7722", "#004b49"), name = "Scale") +
  facet_grid(pop_predict ~ covar_name, scales = "free_x", drop = TRUE) +
  expand_limits(y = 0) +
  xlab("log(Observed bites)") +
  ylab("log(Predicted bites)") +
  labs(tag = "A")

ggplot(data = filter(preds_grouped, intercept == "random"), 
       aes(x = log(avg_bites + 0.1), y = log(bites_mean + 0.1), 
           color = interaction(scale, data_source), shape = ctar_bump)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  scale_color_manual(values = c("darkred", "#cc7722", "#004b49"), name = "Scale") +
  facet_grid(pop_predict ~ covar_name, scales = "free_x", drop = TRUE) +
  expand_limits(y = 0) +
  xlab("log(Observed bites)") +
  ylab("log(Predicted bites)") +
  labs(tag = "B") 
```

### Predictions out-of-fit

#### Moramanga estimates on Mada commune and districts

```{r}
outfit_mora <- read.csv("output/preds/outfit_mora.csv")

ggplot(data = filter(outfit_mora, intercept == "fixed"), 
       aes(x = log(avg_bites + 0.1), y = log(bites_mean + 0.1), 
           color = scale, shape = ctar_bump)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  scale_color_manual(values = c("darkred", "#cc7722"), name = "Scale") +
  facet_grid(pop_predict ~ covar_name, scales = "free_x", drop = TRUE) +
  expand_limits(y = 0) +
  xlab("log(Observed bites)") +
  ylab("log(Predicted bites)") +
  labs(tag = "A")

ggplot(data = filter(outfit_mora, intercept == "random"), 
       aes(x = log(avg_bites + 0.1), y = log(bites_mean + 0.1), 
           color = scale, shape = ctar_bump)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  scale_color_manual(values = c("darkred", "#cc7722"), name = "Scale") +
  facet_grid(pop_predict ~ covar_name, scales = "free_x", drop = TRUE) +
  expand_limits(y = 0) +
  xlab("log(Observed bites)") +
  ylab("log(Predicted bites)") +
  labs(tag = "B")

```

#### Mada commune and districts estimates on Moramanga data
```{r}
outfit_grouped <- read.csv("output/preds/outfit_grouped_mada.csv")

ggplot(data = outfit_grouped, 
       aes(x = log(avg_bites + 0.1), y = log(bites_mean + 0.1), 
           color = scale, shape = ctar_bump)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey") +
  scale_color_manual(values = c("darkred", "#cc7722", "#004b49"), name = "Scale") +
  facet_grid(pop_predict ~ covar_name, scales = "free", drop = TRUE) +
  expand_limits(y = 0) +
  xlab("log(Observed bites)") +
  ylab("log(Predicted bites)")

```

## Burden estimates