### 2019-12-04 supplementary.Rmd
supplementary materials rmd

---
title: "Supplementary Materials"
author: ""
date: 
output:
  html_document: default
  pdf_document:
    includes:
      in_header: format/mystyle.sty
  word_document:
    reference_docx: format/word_styles_reference_01.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## S1. Estimating travel times to the nearest ARMC

```{r ttimes}
```

## S2. Correcting data for underreporting and excluding Category I exposures

### Underreporting of forms
Figure S1.1 correcting
```{r reporting}
```
Figure S1.2 vial ests

### Cat I exposures
Figure S1.3 contacts
```{r cat I}
```
Figure S1.4 ttime differences

## S3. Candidate models of bite incidence

## S4. Sensitivity of these models to reporting/contact cut_offs

## S5. Incrementally adding clinics (where they were added, shifts in ttimes + catchment pops, shifts in burden/reporting etc. --answer q here: why are there still abt 100 deaths)

## S6. Sensitivity of burden estimates to parameter assumptions (baseline + incremental)