### 2019-12-05 manuscript/supplementary.Rmd
.gitignore fix

---
title: "Supplementary Materials"
author: ""
date: 
output:
  word_document:
    reference_docx: format/word_styles_reference_01.docx
  html_document:
  github_document:
  pdf_document:
    includes:
      in_header: format/mystyle.sty
bibliography: ["../manuscript/references.bib"]
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)
opts_chunk$set(echo = FALSE)
```

## S1. Estimating travel times to the nearest ARMC 
To estimate travel times to the nearest clinic at the scale of administrative units, we used two raster inputs: 1) the travel time estimates generated using the friction surface from the Malaria Atlas Project [@weissGlobalMapTravel2018] at an ~1 km^2 scale and the locations of existing ARMC and 2) the population estimates which we resampled from the World Pop ([worldpop.org](https://www.worldpop.org)) raster of population originally at an ~100m^2 resolution. We extracted the mean of travel times (Fig S1.1A) weighted by the population in that grid cell (Fig S1.1B) for each district or commune. Figure S1.1 shows the raster inputs as well as the resulting estimates of travel times at the district (Fig S1.1D) and commune (Fig S1.1D) level.

```{r fig S1.1, fig.fullwidth=TRUE}
include_graphics("../figs/S1.1.jpeg")
```
_**Figure S1.1. A)** Travel time estimates at an approximately 1x1 km scale **B)** Population per 1x1 km grid cell. Mean travel times weighted by population size extracted for each **C)** District and **D)** Commune._

## S2. Correcting data for underreporting and excluding Category I exposures

### Estimating underreporting of forms
Figure S1.1 correcting
```{r reporting}
```
Figure S1.2 vial ests

### Excluding Category I exposures
Figure S1.3 contacts
```{r cat I}
```
Figure S1.4 ttime differences

## S3. Candidate models of bite incidence

## S4. Sensitivity of these models to reporting/contact cut_offs

## S5. Incrementally adding clinics (where they were added, shifts in ttimes + catchment pops, shifts in burden/reporting etc. --answer q here: why are there still abt 100 deaths)

## S6. Sensitivity of burden estimates to parameter assumptions (baseline + incremental)

## References