### 2020-02-18 docs/manuscript.Rmd
updating organization of man

---
title: "How access to care shapes disease burden: the current impact of post-exposure prophylaxis and potential for expanded access to prevent human rabies deaths in Madagascar"
author:
  - Malavika Rajeev:
      email: mrajeev@princeton.edu
      institute: [Princeton]
      correspondence: true
  - Helene Guis:
      institute: [IPMepi, CIRAD]
      correspondence: false
  - Glenn Edosoa:
      institute: [OMS]
      correspondence: false
  - Chantal Hanitriniaina:
      institute: [UoA]
      correspondence: false
  - Soa Fy Andriamandimby:
      institute: [IPMviro]
      correspondence: false
  - Anjasoa Randrianarijaona:
      institute: [IPMepi]
      correspondence: false
  - Reziky Tiandraza Mangahasimbola:
      institute: [IPMepi]
      correspondence: false
  - Ravo Ramiandrasoa:
      institute: [IPMvacc]
      correspondence: false
  - Laurence Baril:
      institute: [IC]
      correspondence: false
  - C.J.E. Metcalf:
      institute: [Princeton]
      correspondence: false
  - Katie Hampson:
      institute: [Glasgow]
      correspondence: false
institute:
  - Princeton: Department of Ecology and Evolutionary Biology, Princeton University, Princeton, United States
  - IPMviro: Virology Unit, Institut Pasteur de Madagascar, Antananarivo, Madagascar
  - IPMvacc: Vaccination Center, Institut Pasteur de Madagascar, Antananarivo, Madagascar
  - IPMepi: Epidemiology and Clinical Research Unit, Institut Pasteur de Madagascar, Antananarivo, Madagascar
  - IC: Institut Pasteur de Cambodge
  - CIRAD: CIRAD, UMR ASTRE, Antananarivo, Madagascar
  - OMS: OMS
  - UoA: Mention Zoologie et Biodiversité Animale, Faculté des Sciences, Université d’Antananarivo, Antananarivo, Madagascar
  - Glasgow: Boyd Orr Centre for Population and Ecosystem Health, Institute of Biodiversity, Animal Health and Comparative Medicine, University of Glasgow, Glasgow, UK
date: ""
link-citations: no
output:
 bookdown::word_document2:
    pandoc_args:
      - --filter=pandoc-crossref
      - --lua-filter=format/scholarly-metadata.lua
      - --lua-filter=format/author-info-blocks.lua
      - --lua-filter=format/pagebreak.lua
      - --lua-filter=format/multiple-bibliographies.lua
      - --reference-doc=format/Reference_Document.docx
 bookdown::html_document2:
    toc: true
    toc_float: true
    pandoc_args:
      - --csl=format/plos-neglected-tropical-diseases.csl
      - --filter=pandoc-crossref
      - --lua-filter=format/scholarly-metadata.lua
      - --lua-filter=format/author-info-blocks.lua
      - --lua-filter=format/pagebreak.lua
      - --lua-filter=format/multiple-bibliographies.lua
keywords: "canine rabies, post-exposure prophylaxis, Gavi, Zero by 30, access to care"
bibliography: ["bib/refs_all.bib"]
# bibliography_main: ["bib/refs_main.bib"]
# bibliography_supp: ["bib/refs_supp.bib"]
---

##### TO DO: fix affiliations

```{r setupmain, include=FALSE}
## Libraries
library(knitr)
library(tidyverse)
library(data.table)
library(rgdal)
library(glue)

opts_chunk$set(echo = FALSE, warning = FALSE)
out_main_fig = TRUE

## data
ctar_metadata <- fread("../data/raw/ctar_metadata.csv")
district_bites <- fread("../output/bites/district_bites.csv")
mora_bites <- fread("../output/bites/mora_bites.csv")
mada_communes <- readOGR("../data/processed/shapefiles/mada_communes.shp")
mada_districts <- readOGR("../data/processed/shapefiles/mada_districts.shp")

```
\newpage

# ABSTRACT
Access to care shapes burden of diseases globally. For dog-mediated rabies, post-exposure prophylaxis (PEP) is highly effective at preventing human deaths, however access to PEP is limited in rabies endemic countries. Recently, a target to eliminate deaths due to canine rabies has been set forth by global health partners. In 2018, Gavi added human rabies vaccine to its investment portfolio, a decision which has the potential to greatly reduce rabies deaths. Here, we estimate rabies burden in Madagascar incorporating estimates of geographic access to PEP. We find that travel times to the closest Anti-Rabies Medical Center (ARMC) is a strong predictor of reported bite incidence across the country. We use resulting estimates of bite incidence in an adapted decision tree framework to generate estimates of burden and reporting across the country, and find that spatial access to PEP shapes local burden. We evaluate the impact of expanding access to PEP, using a strategic framework which minimizes travel time for people at the greatest risk of death given a rabies exposure (persons > 3 hrs away from a clinic). Under these assumptions, we estimate that provisioning of PEP to an additional 200 clinics in Madagascar could reduce the burden of rabies deaths by 50%.  These results highlight the importance of accounting for access to care in predicting the impacts of interventions sub-nationally and the potential for expanding PEP access strategically to reduce rabies deaths in endemic settings.

\newpage

# INTRODUCTION

Inequities in access to care are a major driver of disease burden globally [@world2018world]. Often, the populations at greatest risk of a given disease are the most underserved [@peters2008poverty]. Delivering interventions to these groups is challenging due to financial and infrastructural limitations, and requires careful consideration of how best to allocate limited resources [@mainaSpatialDatabaseHealth2019].

Dog-mediated rabies is estimated to cause approximately 60,000 human deaths annually [@whorabiesmodelingconsortiumPotentialImpactImproved2019]. Mass vaccination of domestic dogs has been demonstrated to be a highly effective way to control the disease in both animals and humans . While dog vaccination can interrput transmission in the reservoir, human deaths can also be prevented through prompt administration of post-exposure prophylactic vaccines (PEP) following a bite by a rabid dog [@RabiesVaccinesWHO2018]. However, access to human vaccines are limited in countries where canine rabies is endemic [@sreenivasanOverviewRabiesPostexposure2019; @tarantolaRabiesPostexposureProphylaxis2017; @hampsonEstimatingGlobalBurden2015], and within countries these deaths are often concentrated in rural, underserved communities [@wentworthSocialJusticePerspective2019].

In 2015, a global framework to eliminate deaths due to dog-mediated rabies by 2030 through a combination of PEP provisioning and dog vaccination was established by the World Health Organization (WHO) and partners [@world2018zero]. Furthermore, in 2018, Gavi, the Vaccine Alliance, added human rabies vaccines to their proposed investment portfolio [@GaviVIS]. From 2021, Gavi-eligible countries will be able to apply for support to expand access to these vaccines, with potential to greatly reduce deaths due to rabies. Yet it is unclear how these vaccines should be distributed to prevent deaths most effectively.

A primary challenge is the lack of data on rabies exposures in humans and incidence in animals in most rabies-endemic countries [@scottAddressingDisconnectEstimated2017]. The majority of rabies burden studies at the national level have integrated bite patient data on reported bites into a probability decision tree framework, with the key assumption that overall reported bite incidence (i.e. both bites due to non-rabid and rabid animals) are proportional to rabies incidence (i.e. the more bites reported in a location, the higher the incidence of *rabies* exposures), and that reporting is uniform across space [@hampsonEstimatingGlobalBurden2015; @knobelReevaluatingBurdenRabies2005; @rajeevHealthcareUtilizationProvisioning2018]. If applied subnationally, these assumptions would likely underestimate rabies deaths in places with poor access to PEP and may overestimate rabies deaths in places with better access to PEP.

In Madagascar, the Institut Pasteur de Madagascar (IPM) provides PEP at no-cost to patients at 31 clinics, or anti-rabies medical centers (ARMC) across the country [@rajeevHealthcareUtilizationProvisioning2018]. PEP is not available at any other public clinics or through the private sector. In addition, there is limited control of rabies in dog populations and the disease appears to be endemic throughout the country [@andriamandimbySurveillanceControlRabies2013; @reynesLaboratorySurveillanceRabies2011]. Due to the spatially restricted nature of PEP distribution points and lack of direct costs for PEP, geographic access is likely to be a major driver of disease burden within the country. Previously, we  estimated the burden of rabies in Madagascar nationally using data from a single district to extrapolate to the country, and did not account for spatial variation in access [@rajeevHealthcareUtilizationProvisioning2018]. Here, we revise estimates of burden by incorporating the impact of access to PEP at the sub-national level on preventing human rabies deaths and explore the potential impact of expanding provisioning of human rabies vaccines on further reducing these deaths.

# METHODS

## Estimating travel time to closest ARMC {#methods-spatial}
We used the global friction surface for 2015 generated by the Malaria Atlas Project (https://map.ox.ac.uk/research-project/accessibility_to_cities/, @weissGlobalMapTravel2018) and GPS points of clinics to estimate the travel time to the nearest ARMC for the country of Madagascar at an approximately 1 x 1 km scale. We calculated a weighted average of travel times by human population at the commune and district levels, using administrative shapefiles from the UN Office for the Coordination of Humanitarian Affairs ([see supplementary section S1](#S1)). To asses the accuracy of this method, we compared this estimate to driving time data collected by IPM during field missions and self-reported travel times from a subset of patients reporting to the Moramanga ARMC. We also compared unweighted average travel times and distance  Human population estimates were taken from the 2015 UN adjusted population projections from World Pop (www.worldpop.org, [cite older WP paper here!]) and aggregated to the commune and district levels. We compared these population estimates to recently released estimates from the 2018 census of Madagascar, and a recent census of three communes in the Moramanga District (see SZ).

## Estimating average annual bite incidence {#methods-bitedata}
We used a database of bite patient forms submitted to IPM from ARMC across the country between 2014 - 2017. These were individual patient data forms that were submitted from clinics as frequently as monthly to annually. These data include details of the administrative district of the bite patient’s address and the date of reporting.

For clinics that did submit data, there was substantial underreporting, with months to even a whole year when forms were not submitted. We therefore estimated clinic level reporting as the proportion of days on which forms were submitted, excluding any periods for which there were no forms submitted for 15 consecutive days ([Figure S2.1A and B](#figS2.1)). To estimate the average annual bites reported for each district, we further excluded data from any years for which there was less than 25% reporting at the clinic level.

The 31 months of data available from the Moramanga district resolved to the commune level (the administrative level below the district, Figure 1B) provide a useful comparison to the national scale data. These data were completely reported with details on the type of exposure documented. We calculated average annual bite incidence to the commune level by taking the monthly average of reported bites and multiplying by 12.

## Modeling reported bite incidence as a function of travel time to closest ARMC {#methods-bitemods}
We modeled incidence of bites as a function of travel time ($T$) as follows: 
$$ \mu = exp(\beta_{t}T \ + \ \beta_0)P$$
where $\mu$ is the expected number of bite patients presenting at the ARMC as a function of travel time to the ARMC, and the human population size, $P$ (an offset which scales the predicted number of bites), for a given source location (district or commune). We fit this model to both the national data resolved to the district level and the Moramanga data resolved to the commune level in a Bayesian framework via MCMC using the R package ‘rjags’. In order to more directly compare estimates between these two datasets, we also modeled the national data with a commune-level travel time covariate ($T_{j}$) so that:

$$\mu = \sum \limits_{j=1}^jexp(\beta_{t}T_j\ + \ \beta_0)P_j$$

As travel times are correlated with human population size (Figure S3.1A), we also compared how well bites were predicted by human population size alone, and in combination with travel time. For the models with population size, we used a model framework with either population size alone ($\mu = exp(\beta_pP + \beta_0)$) or population size and travel times ($\mu = exp(\beta_{t}T + \beta_pP + \beta_0)$) as predictors of bites for the Moramanga data, the national data with covariates at the district level, and the national data with covariates at the commune level.

For the models fitted to the national data, we also attempted to capture variation between clinics by modeling a random catchment effect of the form: $B_{0,k} \sim norm(\mu,\ \sigma)$) where $\mu$ is the mean and $\sigma$ is standard deviation and $B_{0,k}$ is the catchment level intercpet).

To test our models predictions further, we used parameter estimates from each model to generate predictions for data that were not used to fit the model (i.e. we used the estimates from models fitted to the national data to predict the Moramanga data, and used estimates from models fitted to the Moramanga data to predict the national data).

Finally, to check how correcting for underreporting affected our modeling results, we fitted these same models to data uncorrected for reporting (see [supplementary section S4](#S4)).

## Modeling the burden of human rabies deaths sub-nationally {#methods-burden}

<!-- TO DO: simplify discussion here but bring in some figs, etc to Supplementary section to explain assumptions! -->

We estimate deaths due to rabies as a function of the number of bites predicted by our model and estimates of endemic rabies exposure incidence using an adapted decision tree framework (Fig 2). We assume annual rabies incidence of 1% in the dog population in an endemic setting with no mass dog vaccination [@whorabiesmodelingconsortiumPotentialImpactImproved2019]. To generate $E_i$, the human exposure incidence in a given administrative unit (where $i$ is the index of the commune or district) based on the model, we multiply this baseline incidence by the estimated dog population and the exposure rate per rabid dog ($p_{exp}$). We estimate the dog population given a human:dog ratio (HDR) of 7 to generate our maximum expected incidence and an HDR of 35 for our minimum expected incidence, giving a range of 11.1 ($E_{min}$) - 55.7 ($E_{max}$) exposures per 100,000 persons. As there is little data on dog populations in Madagascar [@andriamandimbySurveillanceControlRabies2013], this range encompasses both a wide range of observed HDRs [@knobelReevaluatingBurdenRabies2005; @sambo2018estimating] and generates similar human exposure incidences as observed previously across Africa [cite here!] . $E_i$ is then drawn from a uniform distribution between the minimum and maximum expected number of human exposures for each location.

```{r fig 2, fig.fullwidth=TRUE}
include_graphics("../figs/main/M2.jpeg")
```
*__Figure 2.__Decision tree for burden estimation. For a given administrative unit $i$, human deaths due to rabies ($D_i$) are calculated from model predicted bites ($B_i$). To get $R_i$, the number of reported bites that were rabies exposures, we multiply $B_i$ by $p_{rabid}$, the proportion of reported bites that are rabies exposures. $p_{rabid}$ is constrained such that reported rabies exposures cannot exceed total rabies exposures ($E_i$) and reporting cannot exceed a maximum ($p_{max}$). $E_i$ is generated from assumptions about rabies exposure incidence and the underlying dog population extrapolated from human population sizes (see Table 1, Methods text). $R_i$ is subtracted from $E_i$ to get the number of unreported bites ($U_i$) and then multiplied by the probability of death given a rabies exposure ($p_{infect}$) to get the number of deaths due to rabies ($D_i$). Similarly, deaths averted by PEP are estimated by multiplying $R_i$ by $p_{infect}$, i.e. those who would have died given exposure, but instead received PEP. All parameter values and sources are listed in Table 1.*

$B_i$, the number of total reported bites, is drawn from a poisson distribution with the mean predicted number of bites predicted from the models of incidence as a function of travel times. To get $R_i$, the number of reported rabies exposures, we multiply $B_i$ by $p_{rabid}$, the proportion of reported bites that are rabies exposures. We draw $p_{rabid}$ from a uniform distribution ranging from 0.2 - 0.6 based on investigations of bite patients reporting in the Moramanga District [@rajeevHealthcareUtilizationProvisioning2018]. So that rabid reported bites cannot exceed the total expected number of human rabies exposures or a maximum reporting (as we know that even with minimal travel times, people may not report for PEP for other reasons), we constrain $p_{rabid}$ as follows:
$$ p_{rabid}=\begin{cases} x, & \text{if}\ \frac{E_i\rho_{max}}{B_i} > x \\ \frac{E_i\rho_{max}}{B_i}, &\text{otherwise} \end{cases}$$
where $\rho_{max}$ is the maximum reporting. We estimate $\rho_{max}$ from the Moramanga ARMC data for the commune of Moramanga Ville, the closest commune to the ARMC (average of 3.12 minutes travel time to the clinic), and where contact tracing data estimated that approximately 2% of rabies exposure go unreported [@rajeevHealthcareUtilizationProvisioning2018]. We assume that all rabies exposed patients who report to an ARMC receive and complete PEP, and PEP is completely effective at preventing death due to rabies. 
We subtract $R_i$ from $E_i$ to get the number of unreported exposures ($U_i$). Deaths due to rabies in each administrative unit ($D_i$) are then estimated as:
$$ D_i = binom(U_i,\ p_{death}) $$
Deaths averted are calculated similarly as the reported rabies exposures $R_i$ that would have died in the absence of PEP given $p_{death}$ (i.e $A_i = binom(R_i, \ p_{death})$). Table 1 lists all parameter values and sources used in the decision tree. \

Table:**Table 1.** Parameters used in the decision tree to estimate human rabies deaths at the administrative level.\

| **Parameter**   | **Value**        | **Description**	                 | **Source**|
|:----------------|:----------------:|:----------------------------------|:---------:|
| $p_{exp}$ | 0.39 | Probability of human exposure given a dog rabies case| [@whorabiesmodelingconsortiumPotentialImpactImproved2019] |
| $E_{min} − E_{max}$ |	11.1 - 55.7 per 100,000 persons	| Range of human exposure incidence calculated from an assumed dog rabies incidence of 1%, human:dog ratios between 7 - 35, and $p_{exp}$. | [@whorabiesmodelingconsortiumPotentialImpactImproved2019; @knobelReevaluatingBurdenRabies2005; @sambo2018estimating]|
| $p_{rabid}$	| 0.2 - 0.6 |	Proportion of reported bites that are rabies exposures | [@rajeevHealthcareUtilizationProvisioning2018] |
| $\rho_{max}$ | 0.98	| The maximum reporting possible for any location	Data from Moramanga Ville, Moramanga District | [@rajeevHealthcareUtilizationProvisioning2018] |
| $p_{death}$ |	0.16 | The probability of death given a rabies exposure	| [@changaluchaNeedImproveAccess2018]|

## Estimating the impact of expanding PEP provisioning to additional clinics {#methods-incremental}

```{r ncsbs}
## candidate points
csbs <- read.csv("../data/raw/csbs.csv", stringsAsFactors = FALSE)
csbs %>% 
  filter(type == "CSB2", genre_fs != "Priv", type_fs != "Health Post") %>%
  dplyr::select(CTAR = nom_fs, X_COORD = ycoor, Y_COORD = xcoor) -> csbs
```

We developed a framework to rank clinics by how much they could improve access for people currently underserved (> 3 hours away from any clinic in the country) and to estimate the incremental reduction in burden as PEP is expanded to these clinics. We use the approximate locations of all existing public CSB II (centre de sante niveau II, the level below district hospitals, n = `r nrow(csbs)`) as locations of potential ARMC.

Starting with the current locations, we added one clinic at a time, calculating the proportion of people living < 3 hours away from any clinic for the country. We added the clinic which minimized this metric weighted by the reduction in travel times, and then repeated the process iteratively, ranking clinics and adding the top clinic sequentially. In this iterative process, we also exclude any clinics that once added only reduced travel times below the 3 hour threshold for less than 0.01% of the population (< `r round(sum(mada_districts$pop)*0.0001, -2)` persons). We also explore the scenario in which all CSB II in the country provision PEP to get an estimate of the maximum impact of expanding PEP provisioning. Resulting travel time estimates and locations of when and where clinics were added are shown in [Supplementary section S5](#S5). Finally, we estimate the incremental reduction in burden as each clinic is added using the burden framework described in the previous section.

We estimate how expansion to additional clinics may impact vaccine demand. Specifically, we use our model-predicted estimates of annual bites and then aggregate these to the clinic catchment. As clinics are added, multiple clinics may serve a single district, but in general, the majority of the population within a given commune are closest to a single clinic (Fig S5.3B). Therefore, we assign the catchment clinic as the one which is the closest for the majority of people living in a commune. For the district model, we disaggregate the reported bites to the commune level (assuming uniform incidence within the district). For each catchment, we simulate throughput by randomly assigning patient presentation dates, and then assume perfect compliance (i.e. all patients reporting 3 and 7 days after their first dose) to generate dates of subsequent doses. We use these dates to estimate vial usage given routine vial sharing practices in Madagascar (i.e. each vial is split between two patients, and vials are discarded at the end of each day) and the new abridged intradermal regimen recommended by WHO (2 injection sites on day 0, 3, and 7, [@tarantolaIntradermalRabiesPostexposure2018]). We assume that vial sharing is functioning perfectly, with a background wastage of 0.1 mL per vial, but not accounting for wastage due to other reasons (i.e. wastage due to injection error). We take the mean of 1000 simulations of throughput for each catchment and set of clinics.

## Sensitivity of burden estimates to decision tree parameters {#methods-sensitivity}
We estimate burden across a range of parameter values to test the effects of our model assumptions on estimates of rabies burden. Specifically, we fix rabies incidence at the minimum and maximum of our estimated range, $p_{rabid}$ at 0.2 and 0.6, and $\rho_{max}$ at 0.85 and 0.99 to obtain maximum and minimum estimates of burden. We also examine the impact of systematic variation in rabies incidence with human population (a potential proxy for changes in the dog population, see [supplementary section S6](#S6) for more details) by looking at how estimates of burden change if rabies incidence were to scale positively or negatively with human population size.

## Data availability
All analyses were done in R and using these packages (list packages with commas and citations). All processed data, code, and outputs are archived at this zenodo repository and maintained at this github repo. The raw bite patient data at the national level is maintained by IPM. This database was last queried on *date here* for these analyses. Anonymized raw data from this database are available upon request from IPM. Anonymized raw data collected from the Moramanga District are shared in the archived repository.

## Ethics statement
Data collection from the Moramanga District was approved by the Princeton University IRB (# 7801) and the Madagascar Ministry of Public Health Ethics Committee (# 105-MSANP/CE). Oral informed consent was obtained from all interviewed participants. Data collected from bite patients at the national level are maintained jointly by the Ministry of Health and IPM as a routine part of PEP provisioning.

# RESULTS

## Estimating travel times to and catchment populations of 
Travel times to the closest ARMC varies greatly across the country. We compared these estimates based
on the MAP friction surface to self-reported estimates of hours traveled to the Moramanga clinic for a subset of patients and driving times across the country from IPM. We found that the travel times estimated from the friction surface underestimate true driving times and patient reported travel times to the Moramanga ARMC. Importantly, the friction surface does not account for road condition, traffic, or seasonality in travel times. It also assumes that given multiple types of transport in a cell, the fastest mode of transport takes precedence. From the self reported travel patterns, we know that people use mutliple modes of transport and walking is common adding to the variation (SX). Despite these limitations, the estimated travel times correlate strongly with the reported and groundtruthed driving times. Travel times weighted by population at the grid cell level were a better predicter than unweighted travel times or distance (SY). We use the estimated weighted mean of travel times by population as a proxy for access at the commune/district level in subsequent analyses.

For most districts, the majority of bites were reported to the closest clinic as estimated by our travel time metric (Fig 1A, Fig S1.2). Therefore, we defined the catchment for each clinic as all districts for which the clinic was the closest ARMC for the majority of the population in that district. We excluded any districts in a catchment of a clinic which submitted less than 10 forms (excluded `r nrow(ctar_metadata) - max(district_bites$catch)` catchments, corresponding to `r length(mada_districts) - nrow(district_bites)` districts, in grey in Fig 1A).

```{r fig 1, fig.fullwidth=TRUE}
include_graphics("../figs/main/M1.jpeg")
```
*__Figure 1.A)__ The network of patient presentations at the district level for the national data: circles with black outline represent the total number of patients reporting to each clinic for which we have data. Color corresponds to the clinic catchment. Circles with white outline are the total number of bites reported for that district, and the lines show which ARMC those patients reported to, with the line width proportional to number of patients. Districts in catchments excluded due to lack of forms submitted by the clinic are colored in grey. __B)__ The network of patient presentations at the commune level for the Moramanga data with inset of Madagascar showing the location of the district. Circle with black outline is the location of the Moramanga ARMC with the size proportional to the total number of bites reported to the clinic. __C)__ The estimated average annual bite incidence corrected for reporting for the years 2014 - 2017 (y-axis) for each district plotted against travel time to the closest ARMC (x-axis). Colors correspond to the clinic catchment, shape indicates the dataset, and the size of the point indicates the number of observations (i.e. the number of years for which data was available for the national data; note for Moramanga `r max(mora_bites$nobs, na.rm = TRUE)` months of data were used).*

## Estimating average annual bite incidence at the district and commune level
We found that these corrected estimates of reported bites was in agreement with the vial demand for clinics ([Fig S2.1C](#figS2.1)). See [supplementary section S2](#S2) for more details.

Our final dataset consisted of estimates of average bite incidence for `r nrow(district_bites)` out for `r length(mada_districts)` districts total, and `r nrow(mora_bites)` communes within the catchment of the Moramanga District (Fig 1C). Scale discussion. Outlier discussion.

## Modeling reported bite incidence as a function of travel times {#results-bitemods}
All of the candidate models which included travel times produced similar predictions when compared to the observed data (Fig S3.2). Population size alone was the poorest fit to the data as estimated by DIC, and models with population size as an additional covariate did not generate realistic predictions to the observed data or when used to predict out of fit (Figs S3.2 and S3.3).  Sensitivity to underreproting. For the national data, the models with travel times alone as a covariate and a random catchment generated predictions closest to the observed data (Fig S3.2). In addition, the estimates from the model fitted to the Moramanga data fall into the prediction envelope for the random catchment intercept models (Fig 3). These predictions were robust to how we corrected for underreporting ([supplementary section S4](#S4)). As the random intercept models likely capture the range of possible relationships for any given catchment in our data set, we use estimates from these models to generate subsequent predictions of bite incidence.

```{r fig 3, fig.fullwidth=TRUE}
include_graphics("../figs/main/M3.jpeg")
```
_**Figure 3.** Travel times as a predictor of reported bite incidence per 100,000 persons. The estimated relationship between travel time in hours (x-axis) and mean annual bite incidence per 100,000 persons (y-axis). The lines are the estimates using the mean of the posterior distributions of the parameters for three candidate models: model with travel times at the 1) commune and 2) district level fitted to the national data with a random intercept by catchment, and 3) travel times at the commune level fitted to the Moramanga data with a fixed intercept. The envelopes shows the 95% CI given the random intercept for the commune and district models. The Moramanga data was fitted with a fixed intercept, but falls in the prediction envelope for the commune model with a random intercept. The points show the data used to fit the models (National data (triangles) at the district level used to fit the District and Commune models, and Moramanga data (circles) at the commune level used to fit the Moramanga model)._

## Estimating the burden of human rabies deaths subnationally {#results-burden}

```{r burden results}
natl_burden <- read.csv("../output/preds/natl_burden.csv")
```

Overall, we estimate between `r round(sort(natl_burden$deaths_mean)[1])` - `r round(sort(natl_burden$deaths_mean)[nrow(natl_burden)])` average annual deaths in Madagascar due to rabies depending on the model used (Table 2). We also estimate that under the current system of 31 ARMCs in Madagascar, approximately `r round(sort(natl_burden$averted_mean)[1])` - `r round(sort(natl_burden$averted_mean)[nrow(natl_burden)])` deaths due to rabies are prevented through provisioning of PEP each year. In general, the incidence of rabies deaths increases with travel times to clinics, and there is significant variation sub-nationally at the district level, but also within districts (Figure 4).

*__Table 2.__ Model predictions of mean annual reported bite incidence, total deaths, and total deaths averted at the national level for the two models (commune level and district level models with a random catchment effect); 95% CIs in parentheses.*

```{r Table 2}
natl_burden %>%
  mutate_at(vars(starts_with("bites")), function(x, p) x/p*1e5, p = .$pop) %>%
  mutate_at(vars(bites_mean:averted_lower), round) %>%
  mutate(bites_100k = glue("{bites_mean} ({bites_lower} - {bites_upper})"),
         deaths = glue("{deaths_mean} ({deaths_lower} - {deaths_upper})"),
         averted = glue("{averted_mean} ({averted_lower} - {averted_upper})")) %>%
  select(scale, bites_100k, deaths, averted) -> table1

kable(table1, digits = -0, col.names = c("Model", "Reported bite incidence per 100k", "Burden of deaths", 
                                        "Deaths averted"))           
```

```{r fig 4, fig.fullwidth=TRUE}
include_graphics("../figs/main/M4.jpeg")
```
*__Figure 4.__ Spatial variation in predicted incidence of human rabies deaths per 100, 000 persons. __A)__ for each district (y-axis) in Madagascar, where the diamonds show the predicted incidence for the district model and the squares show the incidence for the commune model for all communes in a district. Points are colored by and districts are ordered by travel times. The vertical lines show the average national incidence of human rabies deaths for the commune (grey) and district (black) models. Incidence mapped to the __B)__ commune level and __C)__ district level from the respective models; grey X’s show locations of current CTAR.*

## The impact of expanding PEP access to additional clinics {#results-incremental}
 We found that expanding PEP to additional clinics reduces rabies deaths, with strategic expansion of access to an additional 100 {inline} clinics reducing burden by almost 50% {inline}. However, benefits accrue less rapidly as ARMC are added, saturating after approximately 200 {inline} additional clinics. As clinics are added, reported bite incidence saturates, and instead there are shifts in which clinics bites are reported to ([see supplementary section S5](#S5)). Even at this maximal PEP access, our model still predicts > 100 {inline} rabies deaths per annum, as average reporting of rabid exposures still remains at approximately 85 % ([Fig S5.4](#figS5.4)). This is not due to saturation of reporting (see distribution of reporting fig in S5), but instead to to broad patterns of remoteness across Mada. Vaccine demand also increases, and more vials are needed per averted death as PEP access is expanded. Our model predictes an increase from 40,000 {inline} vials under current clinic provisioning, to ~50,000 {inline} vials with 200 {inline} clinics, ~60,000 {inline} and approximately 90,000 {inline} vials if all clinics were to provision PEP.

<!-- TO DO: State what happens to throughput + vial sensitivity! -->

```{r fig 5, fig.fullwidth=TRUE}
include_graphics("../figs/main/M5.jpeg")
```

*__Figure 5. A)__ Decrease in deaths due to rabies, __B)__ increase in vials needed per death averted, and __C)__ increase in total # of vials as additional ARMC are added at the national level based on the two models of reported bites. Lines are the mean of 1000 simulations with envelopes representing 95% CIs. The “...” on the axes indicates a break in the axes, and the last point is the burden and vial predictions when all CSBS in the country provision PEP (maximum access).*

## Sensitivity of burden estimates to decision tree parameters {#results-sensitivity}
There is significant uncertainty in the predicted numbers of rabies deaths, but the broad patterns of higher risk with increased travel times are robust for both our estimates of burden and the impacts of expanding access to PEP, even given systematic bias in our assumptions of incidence (see [supplementary section S6](#S6)). Our estimate of the absolute number of deaths and the predicted impact of access is most sensitive to assumptions of underlying rabies exposure incidence in the population. Other parameters don't contribute nearly as much uncertainty. In some ways is it better to present this here?

<!-- TO DO: Univariate sensitivity analyses -->

# DISCUSSION

## Main findings
Travel time to the closest ARMC was a strong predictor of reported bite incidence in Madagascar. Using an adapted decision tree framework and assuming endemic rabies incidence, we find that the burden of rabies is likely to be concentrated in areas with low access to PEP. Even given only 31 clinics, we estimate that current provisioning averts 50% {inline} of deaths that would occur due to rabies in the absence of PEP. We estimate that expanding PEP access has the potential to greatly reduce deaths due to rabies, with expansion to an additional 200 {inline} clinics potentially reducing burden by 50 {inline}. However, our model predicts that these effects saturate, and that even given maximum expansion of PEP (i.e. to all major health centers (n = 1676) in the country), over 100 deaths {inline} may still occur nationally. In addition, vial demand outpaces additional deaths averted, with more vials needed per averted death as clinics are added.

## Limitations
Our modeling framework makes strong assumptions about baseline rabies exposure incidence, assuming a range of 11 - 55 exposures per 100,000 persons in Madagascar. We base these estimates on a reasonable range of human:dog ratios [@knobelReevaluatingBurdenRabies2005; @ratsitorahinaDogEcologyDemography2009; @sambo2018estimating] and dog rabies incidence level [@whorabiesmodelingconsortiumPotentialImpactImproved2019]. A previous study in the Moramanga District in Madagascar estimated a range of 42 – 110 exposures per 100,000 persons, however this was over a 16 month period with frequent rabies cases and multiple human deaths [@rajeevHealthcareUtilizationProvisioning2018]. As incidence of rabies exposures varies significantly over time and space [@hampsonSynchronousCyclesDomestic2007; @hampsonTransmissionDynamicsProspects2009], we sample uniformly over this range and over the proportion of reported bites estimated to be rabid (0.2 - 0.6 from the Moramanga data) to generate a mean estimate of burden. Even when we assume the extremes of these ranges, or that rabies exposure incidence varies systematically with human populations, we find that the qualitative effects of travel times on risk of rabies deaths remains the same. While our sensitivity analyses show that the relative patterns of burden are robust, the magnitude of deaths is most sensitive to our assumption of rabies incidence. Thus, our estimates are more representative of patterns of risk than quantitative differences in burden, and spatial estimates of rabies exposure incidence would be necessary to generate more precise subnational estimates of burden.

The national data we used were underreported and incomplete, but we attempt to correct for these issues (quantifying under-submission of forms and excluding districts from catchments with less than 25% of data reported). Improving data submission and quality from clinics across the country will be important to validate these results, but also to monitor how well clinics are functioning and to develop better provisioning systems. 
<!-- TO DO: SOMETHING ABT SHORTAGES HERE AND BITS FROM CLAIRE'S MSC -->

We make several simplifying assumptions regarding PEP provisioning and efficacy. First, we assume perfect compliance of patients and efficacy of PEP. Although delays in reporting and poor adherence to PEP contribute to risk of death, PEP compliance is generally high in Madagascar [@rajeevHealthcareUtilizationProvisioning2018], and these factors are unlikely to be major drivers of burden [@changaluchaNeedImproveAccess2018]. We also assume that patients report to their closest clinic, but in our data approximately 10% of patients reported to other clinics. As we may not be assigning reported bites to the right clinic, we may potentially be biasing our estimates of vial sharing and throughput at the clinic level. Current practices of provisioning PEP reactively to Category I exposures. Based on data from the Moramanga District, vial demand could be up to 20% lower if these practices were stopped [@rajeevHealthcareUtilizationProvisioning2018].

We assume that spatial access to PEP is the primary driver of health-seeking behavior, however many other factors may drive an individual’s decision to seek PEP, such as socioeconomic and educational status and awareness about the danger of rabies [@samboKnowledgeAttitudesPractices2014; @barbosacostaInfluencePovertyRabies2018; @etheartEffectCounsellingHealthcareseeking2017; @hampsonRabiesExposuresPostExposure2008]. Importantly, in the Madagascar context where PEP is provided free of charge to patients, the main cost of seeking PEP is for transport and associated time lost for travel. In addition, more remote communities tend to be of lower socioecomonic and educational status [@peters2008poverty], so our travel time metric may act as a proxy for these set of correlated variables that interact to determine an individual’s access to PEP. While we did not explicitly include variation in clinic functioning, we did include a catchment level random intercept, to account for spatial differences in reported bite incidence not due to travel times. This random component of the model did improve fit to the data, suggesting that variation at the clinic level (i.e. differences in clinic functioning such as vaccine availability) and at the catchment level (i.e. differences in regional bite incidence, awareness, outreach, etc.) are likely important drivers of health-seeking behavior and overall vaccine demand.

## Broader context
```{r}
burden_preds <- fread("../output/preds/complete/burden_filled.csv")
## Grouped to district
burden_preds$distcode <- mada_communes$distcode[burden_preds$names]
burden_preds$commcode <- mada_communes$commcode[burden_preds$names]

burden_preds %>%
  filter(distcode == "MG33314", scenario == 0) %>%
  group_by(scale) %>%
  summarise(death_inc = sum(deaths_mean)/sum(pop)*1e5,
            reporting = mean(reporting_mean)) -> mora_data

```
While our incidence of human rabies deaths are based on strong assumptions of rabies incidence, they generally fall within the range estimated for a wide range of settings (1 - 5 subnationally, on average 1.5, see [@beyeneEstimatingBurdenRabies2018; @hampsonRabiesExposuresPostExposure2008; @sudarshanAssessingBurdenHuman2007; @tenzinDogBitesHumans2011; @lyRabiesSituationCambodia2009]. In particular, the estimate of incidence of deaths for the Moramanga District fell within the range of incidence estimated from a contact tracing study in this district (1.35 {inline} vs. 1-3 deaths {inline} per 100,000 persons and reporting 72 vs 85 {inline}, [@rajeevHealthcareUtilizationProvisioning2018]). At the national level, they are in line with estimates of burden based on a more standard decision tree approach [@whorabiesmodelingconsortiumPotentialImpactImproved2019].

As bites due to other reasons than rabies can be more common than genuine rabies exposures [@rysavaPathRabiesElimination2018; @wallaceEstablishmentCanineRabies2015], reported bite data may be a poor indicator of rabies transmission, particularly in settings where barriers to PEP are low (i.e low or no direct costs, low travel times, high socioeconomic status of the population). Surveillance instead...Integrated bite case management and clinic-based surveillance approaches have been established as feasible tools that could accomplish this [@lechenneImportanceParticipatoryIntegrated2017; @undurragaCostEffectivenessEvaluationNovel2017], and ARMC could act as sentinels for rabies surveillance across the country. Ultimately, these methods rely on exposed persons to report to clinics, and as such, expanding access to PEP could strengthen surveillance.

Our work is in line with previous studies demonstrating the impacts of geographic access to care on health outcomes [@aleganaTreatmentseekingBehaviourLow2017; @nikolayEvaluatingHospitalBasedSurveillance2017; @macphersonDisparitiesAccessDiagnosis2019; @polettiHiddenBurdenMeasles2018; @Hegde2019dt; @metcalfTransportNetworksInequities2015]. While many of these studies directly estimated reporting using community based and cross-sectional survey data [@aleganaTreatmentseekingBehaviourLow2017; @nikolayEvaluatingHospitalBasedSurveillance2017], we use model derived estimates of reporting and assumptions of baseline disease incidence to estimate burden, similar to methods developed for other pathogens to estimate burden in data sparse settings [@macphersonDisparitiesAccessDiagnosis2019; @polettiHiddenBurdenMeasles2018; @Hegde2019dt]. For canine rabies, estimating burden of disease directly is challenging [@scottAddressingDisconnectEstimated2017; @taylorDifficultiesEstimatingHuman2017] and direct empirical estimates of rabies burden have previously relied on exhaustive contact tracing studies [@beyeneEstimatingBurdenRabies2018; @hampsonRabiesExposuresPostExposure2008] or large-scale mortality surveys [@sudarshanAssessingBurdenHuman2007]. Recent studies have estimated access to health-seeking behavior and PEP completion and adherence, but not directly linked these estimates to burden [@tarantolaRabiesPostexposureProphylaxis2017; @zaidiGeographicVariationAccess2013]. Our framework for incorporating estimates of access into previously developed burden estimation methods (i.e. decision trees) could help to improve and refine existing burden estimates and explore impacts of expanding access, particularly at the sub-national level.

In most rabies endemic countries, access to PEP is geographically limited and/or costly to patients [@changaluchaNeedImproveAccess2018; @liDescriptiveAssessmentRabies2019; @sreenivasanOverviewRabiesPostexposure2019; @tarantolaRabiesPostexposureProphylaxis2017].  Currently, each clinic in Madagascar serves on average a catchment population of 750,000 persons {inline}, compared to Tanzania where catchment sizes are about half of that. Latin American countries, where significant progress has been made towards elimination, aim to provide one clinic providing PEP for 100,000 persons. Previous to 2006 when the nerve tissue vaccine was still in use, access to PEP in Madagascar was more widespread, with clinics able to request vaccine upon demand. However, once the more Vero cell vaccine was implemented, vaccine provisioning to all clinics became too costly, resulting in provisioning to a subset of clinics [@andriamandimbySurveillanceControlRabies2013]. However, Gavi investment can remove this cost barrier. If this results in expansion to underserved areas and strengthens existing provisioning and supply chain mechanisms, then rabies burden could be greatly reduced. Given existing challenges in data collection and monitoring of how and to whom PEP is provisioned, strong monitoring and evaluation systems will be necessary to ensure these vaccines are being used judiciously and equitably.

Ultimately even given expanded access, achieving 'the last mile', i.e. preventing deaths due to rabies in the most remote populations will require a disproportionate level of resources and may not be achievable through PEP alone [@delriovilasTribulationsLastMile2017]. IBCM programs and risked based assessments are one way to tackle the issue of managing PEP resources effectively even as demand outpaces reductions in burden [@rysavaPathRabiesElimination2018]. Combining PEP provisioning with mass dog vaccination will be key to interrrupting transmission, and eliminating deaths due to rabies. These same issues of access, however apply to both dog vaccination programs and surveillance systems, and integrating spatial heterogeneities in vaccination and detection will be critical to understand how control and prevention interventions can be best implemented [@fergusonHeterogeneitySpreadControl2015].

## Conclusion {inline these stats}
Geographic access to PEP is a strong predictor of reported bite incidence in Madagascar and expanding PEP availability to more clinics across the country could greatly reduce human rabies deaths. Based on our modeling framework, a more equitable target may be aiming to minimize the population living within 3-5 hours travel time of a clinic, but due to broad patterns of remoteness across Madagascar, even when all primary health facilities are providing PEP, approximately 10 % {inline} of people live > 3 hours away from a clinic. Expanding PEP to clinics that target currently underserved populations may be the best strategy for preventing human rabies deaths. If expanded only to one clinic per district (N = 114), an additional 40% of deaths could be averted. Expansion of PEP can greatly reduce risk and inequities in access to this life-saving vaccine, but ultimately, interrupting transmission of rabies through mass dog vaccination programs will be the most equitable way to prevent, and eventually eliminate, rabies deaths.

## Acknowledgements

\newpage

## References
::: {#refs_main}
:::

\newpage

# SUPPLEMENTARY MATERIALS AND METHODS

## S1. Estimating travel times to the closest ARMC {#S1}
[*jump back to main methods section: Spatial covariate data*](#methods-spatial)  
To estimate travel times to the nearest clinic at the scale of administrative units, we generated two raster inputs: 1) the travel time estimates generated using the friction surface from the Malaria Atlas Project [@weissGlobalMapTravel2018] at an ~1 km^2 scale and the locations of existing ARMC and 2) the population estimates which we resampled from the World Pop ([worldpop.org](https://www.worldpop.org)) raster of population originally at an ~100m^2 resolution. We extracted the mean of travel times (Fig S1.1A) weighted by the population in that grid cell (Fig S1.1B) for each district (Fig S1.1C) or commune (Fig S1.1D).

(euclidean distances for the driving data with GPS point locations and distance from the centroid of commune where the patient resided to the location of the ARMC for the self reported travel times)

```{r fig S1.1, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S1.1.jpeg")
```
*__[Figure S1.1.]{#figS1.1} A)__ Travel time estimates at an approximately 1x1 km scale __B)__ Population at same ~1x1 km scale. Mean travel times weighted by population size extracted for each __C)__ District and __D)__ Commune.*

We assigned clinic catchments by determining which clinic was closest in terms of travel times for the majority of the population within the administrative unit. Most districts and communes a single clinic served the majority of the population by this metric (Fig S1.2A). In addition, our data show that the majority of bite patients do report to the closest clinic as assigned by our travel time metric (Fig S1.2B).

```{r fig S1.2, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S1.2.jpeg")
```
*__[Figure S1.2.]{#figS1.1} A)__ Distribution of the proportion of the population in a given administrative unit (district or commune) served by the catchment assigned. __B)__ The proportion of bites reported to each clinic which originated from a district within the assigned catchment.*

## S2. Estimating average annual bite incidence at the district and commune level
[*jump back to main methods section: Bite patient data*](#methods-bitedata)

To correct for under-submission of forms, we attempted to identify what proportion of data were missing each year by excluding any periods of time for which there were 15 consecutive days where zero patient records were submitted ([Figure S2.1A](figS2.1). We validated this cut-off of consecutive days by comparing estimated vial demand given the total reported bites (corrected for under-submission) for the years 2014-2016 to actual vials provisioned to clinics over this same period. We did not have the date or month when vials were provisioned, so we compared the totals across this three year period.

Vial demand was simulated under simplified assumptions of PEP administration and adherence[@rajeevHealthcareUtilizationProvisioning2018], assuming that patients reported randomly across the year and then completed doses 3, 7, and 28 days subsequent from the date reported. For this analysis, we look at the range of vial estimates assuming that all patients complete 3 vs. 4 doses total. During this period, the Thai Red Cross Intradermal regimen was used across Madagascar, with 0.2 mL administered per patient at each visit. Vials can be shared within a day between two patients, resulting in 0.1 background wastage per vial shared + any additional wasteage from unused doses discarded at the end of the day. While these simplifying assumptions do not capture variation in adherence, completion, delays, or errors in administration, PEP completion is overall high in Madagascar and in other settings where PEP is provided at no direct costs to patients, so our estimates at 3 doses per patient vs. 4 are likely encompassing much of this variation at the annual level.

We find that estimates of vial demand based on uncorrected bite patient numbers are generally lower than the true number of vials provisioned for those clinics with substantial underreporting ([Figure S2.1B](#figS2.1)). Correcting patient numbers by under-submission of forms based on the 15 day consecutive day cut-off results in estimates of vial demand closer to the observed vials provisioned.

```{r fig 2.1, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S2.1.jpeg")
```
*__[Figure S2.1.]{#figS2.1}A)__ The daily time series of the number of forms submitted by each clinic, with periods of time where no forms were submitted for >= 15 days excluded (in white). __B)__ Reporting estimates for each clinic (points are the estimate for each year) calculated as the # of days in a year which were not excluded based on the  criteria of 15 consecutive days of non-reporting/365. __C)__ The difference between log(estimated) and log(observed) vials provisioned for the period of 2013-2016 for each clinic correcting for underreporting (squares) using the 15 day cut-off show in A and B, vs. not correcting for underreporting (circles). We did not have data on vials provisioned for IPM.*

## S3. Modeling reported bite incidence as a function of travel times
<!--  TO DO = COMBINE SECTIONS S3/S4 -->
### Candidate models of bite incidence {#S3}

[*jump back to main results section: Travel times as a predictor of bites*](#results-bitemods)

We compared alternative model structures for predicting reported bites, specifically by including human population size as a predictor of reported bites (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model), as well as by intercept type ( random = random intercept by catchment, fixed = a single fixed intercept was estimated) for the national data (see main Methods for specific formulations). We used priors centered around zero (*norm(0, 10-6)*) except for the precision parameter for which we used an uniform distribution (*unif(0, 100)*).\

The Moramanga data were not fit to models with a random catchment effect, as these were data from a single catchment. We calculated Deviance Information Criterion (DIC, a metric of model fit to data) for each candidate model as well as potential scale reduction factors (psrf, a metric of model convergence) for each covariate in the model and the multivariate psrf for the whole model (cite Gelman book). All models and posterior estimates of parameters converged (Fig S3.1, both psrf and mpsrf approach 1). Population size was the poorest predictor of bites (Fig S3.2, Table S3.1) and adding population size to the model did not improve the fit to the national level data (Figure S3.2), although these models did have lower DIC values for the Moramanga data (Table S3.1). However, when used to predict data that were not used to fit the models (i.e. model estimates from fitting to the Moramanga data were used to predict the national data and vice versa), only models with incidence offset by population size (flatPop) rather than scaling with population size (addPop) generated reasonable predictions to the data (Figure S3.3).\

*__Table S3.1.__ DIC Estimates for all models. Models with the lowest DIC for each dataset and scale of covariate are in bold.  For the column pop effect, addPop = models with population size as additional covariate, onlyPop = models with population as only covariate, flatPop = models with population as offset in model. For the intercept type: random = random intercept by catchment, fixed = a single fixed intercept was estimated).*
```{r table 1}
model_ests <- read.csv("../output/mods/estimates.csv")
model_ests %>%
  select(params, Mean, pop_predict, intercept, data_source, scale) %>%
  spread(key = params, value = Mean, fill = 0) -> model_means
model_ests %>%
  group_by(data_source, scale, pop_predict, intercept) %>%
  summarize(dic = mean(dic)) -> dic_ests
dic_ests %>%
  group_by(data_source) %>%
  arrange(dic, .by_group = TRUE) %>%
  mutate(dic = round(dic, 2)) -> dic_ranks
knitr::kable(dic_ranks, 
             col.names = c("Dataset", "Scale", "Pop effect", "Intercept type", "DIC"))
```
<br/>

```{r fig S3.1, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S3.1.jpeg")
```
_**Figure S3.1. Potential scale reduction factors (psrf) for all models.** The x-axis is the type of psrf, either multivariate which assesses the convergence of the model, or the individual which assesses the convergence for each parameter estimated from the model. Columns are by the type of model intercept (either a fixed intercept or a random intercept by catchment) and rows are the type of model structure with respect to the population covariate (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model). Colors show which data set was used for fitting and the scale of the model (Moramanga = Moramanga data with covariates at the commune level, Commune = National data with covariates at the commune level, District = National data with covariates at the district level)._\  
<br/>

```{r fig S3.2, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S3.2.jpeg")
```
_**Figure S3.2. Prediction to data used to fit each model.** Log of the observed bites against the predicted bites using the mean of the posterior distributions for each parameter. Columns are by the type of model intercept (either a fixed intercept or a random intercept by catchment) and rows are the type of model structure with respect to the population covariate (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model). Colors show which data set  was used for fitting and the scale of the model (Moramanga = Moramanga data with covariates at the commune level, Commune = National data with covariates at the commune level, District = National data with covariates at the district level)._\

<!-- Example of how to link to full figure online -->
<!-- [link to full fig](https://raw.githubusercontent.com/mrajeev08/ModelingChapter/master/figs/fig1.jpeg) -->

<br/>

```{r fig S3.3, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S3.3.jpeg")
```
_**Figure S3.3. Out of fit predictions to data.** Log of the observed bites against the log of predicted bites generated from the mean of the posterior distributions from each model for data not used to fit the model. First two columns are the predictions from the commune and district model fitted to the national data for the Moramanga data with fixed and random intercepts. The third column are predictions from models fitted to the Moramanga data predicting the national data at the commune and district scale (only fixed intercept models). Rows are the type of model structure with respect to the population covariate (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model). Colors show which data set was used for fitting and the scale of the model (Moramanga = Moramanga data with covariates at the commune level, Commune = National data with covariates at the commune level, District = National data with covariates at the district level)._

<br/>

### Sensitivity of models to correction for underreporting {#S4}

<!-- TO DO: add one for lower cut-off (i.e. 5 days consec without reporting) -->
[*jump back to main results section: Models*](#results-bitemods)

We also looked at the sensitivity of our model estimates to how we corrected our data for underreporting. Specifically we compared results from models fitted x, y, x.

We found that the model with population size as an offset (flatPop) still generated the closest predictions to the data (Table S4.1 for DIC comparisons and Fig S4.1 for predictions to the data). These models with a random intercept also resulted in qualitatively similar predictions of the impact of travel times on reported bite incidence, although correcting for underreporting did impact the intercept (modulating the overall incidence estimates, but not the relationship with travel times, Fig S4.2).

*__Table S4.1. Comparing DIC ranks for models using datasets with different data correction. For the column pop effect, addPop = models with population size as additional covariate, onlyPop = models with population as only covariate, flatPop = models with population as offset in model. For the intercept type: random = random intercept by catchment, fixed = a single fixed intercept was estimated).*

```{r Table S4.1}
## DIC Tables
model_se <- fread("../output/sensitivity/model_se.csv")
model_se %>%
  group_by(scale, contacts, reporting, intercept, pop_predict) %>%
  summarize(dic = dic[1]) %>%
  arrange(reporting, contacts, dic) %>%
  mutate(colnames = interaction(reporting, contacts),
         colnames = case_when(colnames == "15.3" ~ "Corrected for underreporting & Cat I excluded",
                              colnames == "Inf.3" ~ "Cat I excluded only",
                              colnames == "15.Inf" ~ "Corrected for underreporting only",
                              colnames == "Inf.Inf" ~ "None (raw data)")) %>%
  ungroup() %>%
  select(-contacts, - reporting) %>%
  mutate(dic = round(dic, 2)) %>%
  pivot_wider(names_from = colnames, values_from = dic) -> dic_table

knitr::kable(dic_table,
             col.names = c("Scale", "Intercept type", "Pop effect", "Corrected for underreporting & Cat I excluded", "Corrected for underreporting only", "Cat I excluded only", "None (raw data)"))
```

```{r fig S4.1, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S4.1.jpeg")
```
_**Figure S4.1 Predictions to the national data with different corrections.** Log of the observed bites against the log of predicted bites generated from the mean of the posterior distributions from each model for data corrected for both underreporting and with Category I exposures excluded, only corrected for underreporting, with only Category I exposures excluded, and the raw data with no corrections. Rows are the type of model structure with respect to the population covariate (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model) and the intercept type (random intercept by catchment or fixed intercept). Colors show which data set was used for fitting and the scale of the model (Moramanga = Moramanga data with covariates at the commune level, Commune = National data with covariates at the commune level, District = National data with covariates at the district level)._

```{r fig S4.2, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S4.2.jpeg")
```
_**Figure S4.2.** The estimated relationship between travel time in hours (x-axis) and mean annual bite incidence per 100,000 persons (y-axis) for models with a random catchment intercept and with population as an offset. Panels show predictions from models fitted to (1) data corrected for both underreporting and with Category I exposures excluded, (2) only corrected for underreporting, (3) with only Category I exposures excluded, and the raw (4) data with no corrections (raw data). The lines are the estimates using the mean of the posterior distributions of the parameters and the envelopes shows the 95% CI given the random intercept. The colors are the two candidate models with travel times at the commune and district level fitted to the national data with a random intercept by catchment. The points show the data used to fit the models._

## S5. Estimating the impact of expanding PEP provisioning to additional clinics {#S5}
[*jump back to main methods section: Incremental reduction of burden*](#methods-incremental)  
[*jump back to main results section: Incremental reduction of burden*](#results-incremental)

```{r scenario stats}
## candidate points
csbs <- read.csv("../data/raw/csbs.csv", stringsAsFactors = FALSE)
csbs %>% 
  filter(type == "CSB2", genre_fs != "Priv", type_fs != "Health Post") %>%
  dplyr::select(CTAR = nom_fs, X_COORD = ycoor, Y_COORD = xcoor) -> csbs
catches_incremental <- fread("../output/stats/pop_by_catch.csv")
```

We algorithimically added clinics by their relative reductions in travel times for people living > 3 hours away from a clinic. Fig S5.1 shows the location and at what step clinics were added. `r max(catches_incremental$scenario[catches_incremental$scenario != max(catches_incremental$scenario)])` clinics were added before there were no clinics which reduced travel times for more than 0.01% of people who also lived > 3 hours away from a clinic. We also estimated travel times given that all additional CSB IIs provisioned PEP (an additional `r nrow(csbs) - 31` clinics). Fig S5.2 shows the changes in travel times at the grid cell and administrative unit scale as clinics are added. As clinics were added, more than one clinic could serve a single district (the proportion of the pop fig S5.3). Even at the maximum level of access, 10% of people still live more than 3 hours from a clinic hours (S5.4). Fig S.55 shows how average travel times, bite incidence, reporting of rabid exposures, and incidence of deaths at the commune and district scale decline as clinics are added. Fig S56 shows the shifts in these distribution as clinics are added. At the clinic level, Fig S5.7 shows the shifts in the distribution of catchment populations, annual vial demand, and average daily throughput as clinics are added (S5.8).

```{r fig S5.1, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.1.jpeg")
```
_**Figure S5.1.** Maps of where and at what step (i.e. commune/district where the Nth clinic was added) at the **A)** district and **B)** commune level. X's indicate the location of existing ARMC (N = 31), polygons in white are locations where an additional clinic was not added based on the algorithm criteria (reducing travel times for > 0.01 % of people living > 3 hours away from any clinic)._

```{r fig S5.2, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.2.tiff")
```
_**Figure S5.2.** Maps of how travel times change as clinics are added at the **A)** ~1 x 1 km grid cell **B)** commune and **C)** district scales. The columns are ordered by the number or clinics at each step: baseline (N = 31), 100, 200, 300, 472 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country). Grey pixels show the location of ARMC at each step. Commune and district values are the average grid cell travel times weighted by the population in each cell._

```{r fig S5.3, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.3.jpeg")
```
_**Figure S5.3.** As clinics are added, shifts in distribution of the proportion of population in a given catchment at the commune and district scales (colors). Rows are ordered by the number or clinics at each step: baseline (N = 31), 100, 200, 300, 472 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country)._

```{r fig S5.4, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.4.jpeg")
```
_**Figure S5.4.** As clinics are added, the proportion of the population binned by travel times in hours. Panel labels are the number or clinics at each step: baseline (N = 31), 100, 200, 300, 472 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country)._

```{r fig S5.5, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.5.jpeg")
```
_**Figure S5.5.** As clinics are added, declines in commune and district averages of **A)** travel times (hrs), **B)** bite incidence per 100,000 persons, **C)** reporting of rabid exposures, and **D)** incidence of deaths per 100,000 persons._

```{r fig S5.6, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.6.jpeg")
```
_**Figure S5.5.** As clinics are added, shifts in the distribution of **A)** travel times (hrs), **B)** bite incidence per 100,000 persons, **C)** reporting of rabid exposures, and **D)** incidence of deaths per 100,000 persons at the commune and district scales (colors). Rows are ordered by the number or clinics at each step: baseline (N = 31), 100, 200, 300, 472 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country)._

```{r fig S5.7, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S5.7.jpeg")
```
_**Figure S5.7.** As clinics are added, declines in clinic average **A)** catchment population size, **B)** daily throughput, and **C)** annual vial demand given estimates of bite incidence for the two models (commune and district, colors)._

## S6. Sensitivity of burden estimates to parameter assumptions {#S6}

[*jump back to main methods section: Incremental reduction of burden*](#methods-sensitivity)  
[*or to main results section: Incremental reduction of burden*](#results-sensitivity)

## SN. Assumptions of underlying rabies incidence 
<!-- TO DO: HERE OR EARLIER BIT? -->

We explore the sensitivity of our burden estimates and estimates of the impacts of PEP expansion to assumptions of rabies exposure incidence. Specifically, we estimate burden fixed at the minimum (11.1 exposures per 100k persons) and maximum (55.7 exposures per 100k persons)  exposure incidence within our estimated range. We also generate scaling factors to look at how systematic variation in rabies incidence with human population size could affect our estimates of burden (Fig S6.1). Specifically, if we believe that human to dog ratios are positively correlated with human populations (i.e. if dog ownership/populations are higher in more populated, urban areas), then we might expect higher rabies exposure incidence within our range as population size increases. Alternatively, if we believe that the human to dog ratio is inversely correlated with human population size (i.e. if dog ownership is more common in less populated, rural areas), we would expect rabies exposure incidence within our range to scale negatively with human population size.

```{r fig S6.1, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S6.1.jpeg")
```
_**Figure S6.1.**Range of constrained scaling factors generate for district and commune population size. The black dashed lines represent the minimum and maximum estimated rabies exposure incidence (minimum = 11.1 exposures/100k persons, maximum = 55.7 exposures/100k persons). We generate constrained scaling factors with human population size so that rabies exposure incidence in each location scales positively or negatively across the range of observed population sizes (panels)._

Overall, our sensitivity analyses show that while there is significant uncertainty in the absolute numbers of deaths due to rabies, patterns with increasing risk of death with reduced access and the potential for expanding access to PEP to reduce these deaths appear to be robust (Figs S6.2, S6.3). In general, for a given incidence level the higher our assumption of the proportion of reported bites which are rabid, the lower the estimated burden. The lower our estimate of maximum reporting, the higher overall burden of rabies, and as this parameter decreases, the relationship between travel times and rabies deaths are dampened. Scaling with population size modulates this impact of travel times on deaths, with positive scaling of rabies incidence with population size (i.e. more rabies in more populated places) dampening and negative scaling enhancing the relationship between access and deaths. Finally, raw numbers of deaths greatly depend on what we assume the rabies exposure incidence is in the population.

```{r fig S6.2, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S6.2.jpeg")
```
*__Figure S6.2.__Predicted relationship between travel times and rabies death incidence per 100k persons for the district model vs. commune models (colors) with columns arranged by the scaling factor (minimum incidence of 11.1 rabies exposures/100,000 persons, the maximum negative scaling with population size, the maximum positive scaling with population size, and maximum incidence of 55.7 rabies exposures/100,000 persons, see Fig S6.1 for the scaling factors). Rows show burden estimates given the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]). Colors represent the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]); shapes show the maximum and minimum assumption of $\rho_{max}$ = [0.8 , 0.95]).*

```{r fig S6.3, fig.fullwidth=TRUE}
include_graphics("../figs/supplementary/S6.3.jpeg")
```
*__Figure S6.3.__ Estimates of the impact of expanding access to PEP incrementally on  burden at the national level for the district model vs. commune models (colors) with columns arranged by the scaling factor (minimum incidence of 11.1 rabies exposures/100,000 persons, the maximum negative scaling with population size, the maximum positive scaling with population size, and maximum incidence of 55.7 rabies exposures/100,000 persons--see Fig S6.1 for the scaling factors). Rows show burden estimates given the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]). Colors represent the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]); shapes show the maximum and minimum assumption of $\rho_{max}$ = [0.8 , 0.95].*

## Supplementary references
::: {#refs_supp}
:::