### 2020-04-19 docs/manuscript.Rmd
WIP miscellaneous edits
- adding in out sessions to some figs
- sub commands to others
- work in progress!

---
title: "How access to care shapes disease burden: the current impact of post-exposure prophylaxis and potential for expanded access to prevent human rabies deaths in Madagascar"
author:
  - Malavika Rajeev:
      email: mrajeev@princeton.edu
      institute: [Princeton]
      correspondence: true
  - Helene Guis:
      institute: [IPMepi, CIRAD]
      correspondence: false
  - Glenn Edosoa:
      institute: [OMS]
      correspondence: false
  - Chantal Hanitriniaina:
      institute: [UoA]
      correspondence: false
  - Soa Fy Andriamandimby:
      institute: [IPMviro]
      correspondence: false
  - Anjasoa Randrianarijaona:
      institute: [IPMepi]
      correspondence: false
  - Reziky Tiandraza Mangahasimbola:
      institute: [IPMepi]
      correspondence: false
  - Ravo Ramiandrasoa:
      institute: [IPMvacc]
      correspondence: false
  - Laurence Baril:
      institute: [IC]
      correspondence: false
  - C.J.E. Metcalf:
      institute: [Princeton]
      correspondence: false
  - Katie Hampson:
      institute: [Glasgow]
      correspondence: false
institute:
  - Princeton: Department of Ecology and Evolutionary Biology, Princeton University, Princeton, United States
  - IPMviro: Virology Unit, Institut Pasteur de Madagascar, Antananarivo, Madagascar
  - IPMvacc: Vaccination Center, Institut Pasteur de Madagascar, Antananarivo, Madagascar
  - IPMepi: Epidemiology and Clinical Research Unit, Institut Pasteur de Madagascar, Antananarivo, Madagascar
  - IC: Institut Pasteur de Cambodge
  - CIRAD: CIRAD, UMR ASTRE, Antananarivo, Madagascar
  - OMS: OMS
  - UoA: Mention Zoologie et Biodiversité Animale, Faculté des Sciences, Université d’Antananarivo, Antananarivo, Madagascar
  - Glasgow: Boyd Orr Centre for Population and Ecosystem Health, Institute of Biodiversity, Animal Health and Comparative Medicine, University of Glasgow, Glasgow, UK
date: ""
link-citations: no
output:
 bookdown::word_document2:
    pandoc_args:
      - --filter=pandoc-crossref
      - --lua-filter=format/scholarly-metadata.lua
      - --lua-filter=format/author-info-blocks.lua
      - --lua-filter=format/pagebreak.lua
      - --lua-filter=format/multiple-bibliographies.lua
      - --reference-doc=format/Reference_Document.docx
 bookdown::html_document2:
    toc: true
    toc_float: true
    pandoc_args:
      - --csl=format/plos-neglected-tropical-diseases.csl
      - --filter=pandoc-crossref
      - --lua-filter=format/scholarly-metadata.lua
      - --lua-filter=format/author-info-blocks.lua
      - --lua-filter=format/pagebreak.lua
      - --lua-filter=format/multiple-bibliographies.lua
keywords: "canine rabies, post-exposure prophylaxis, Gavi, Zero by 30, access to care"
bibliography: ["bib/refs_all.bib", "bib/refs_pkgs.bib"]
# bibliography_main: ["bib/refs_all.bib"]
# bibliography_supp: ["bib/refs_pkgs.bib"]
---

##### TO DO: fix affiliations

```{r setupmain, include=FALSE}
## Libraries
library(knitr)
library(tidyverse)
library(data.table)
library(rgdal)
library(glue)
library(here)
library(magrittr)

opts_chunk$set(echo = FALSE, warning = FALSE)

## data
ctar_metadata <- fread(here("data/raw/ctar_metadata.csv"))
district_bites <- fread(here("output/bites/district_bites.csv"))
mora_bites <- fread(here("output/bites/mora_bites.csv"))
mada_communes <- readOGR(here("data/processed/shapefiles/mada_communes.shp"))
mada_districts <- readOGR(here("data/processed/shapefiles/mada_districts.shp"))

trans_perc <- function(x) {
  round(x, 2)*100
}

```
\newpage

# ABSTRACT
Access to care shapes burden of diseases globally. For dog-mediated rabies, post-exposure prophylaxis (PEP) is highly effective at preventing human deaths, however access to PEP is limited in rabies endemic countries. Recently, a target to eliminate deaths due to canine rabies has been set forth by global health partners. In 2018, Gavi added human rabies vaccine to its investment portfolio, a decision which has the potential to greatly reduce rabies deaths. Here, we estimate rabies burden in Madagascar incorporating estimates of geographic access to PEP. We find that travel times to the closest Anti-Rabies Medical Center (ARMC) is a strong predictor of reported bite incidence across the country. We use resulting estimates of bite incidence in an adapted decision tree framework to generate estimates of burden and reporting across the country, and find that spatial access to PEP shapes local burden. We evaluate the impact of expanding access to PEP, using a strategic framework which minimizes travel time for people at the greatest risk of death given a rabies exposure (persons > 3 hrs away from a clinic). Under these assumptions, we estimate that provisioning of PEP to an additional 200 clinics in Madagascar could reduce the burden of rabies deaths by 50%. These results highlight the importance of accounting for access to care in predicting the impacts of interventions sub-nationally and the potential for expanding PEP access strategically to reduce rabies deaths in endemic settings.

# AUTHOR SUMMARY

\newpage

# INTRODUCTION

Inequities in access to care are a major driver of disease burden globally [@world2018world]. Often, the populations at greatest risk of a given disease are the most underserved [@peters2008poverty]. Delivering interventions to these groups is challenging due to financial and infrastructural limitations, and requires careful consideration of how best to allocate limited resources [@mainaSpatialDatabaseHealth2019].

Dog-mediated rabies is estimated to cause approximately 60,000 human deaths annually [@whorabiesmodelingconsortiumPotentialImpactImproved2019]. Mass vaccination of domestic dogs has been demonstrated to be a highly effective way to control the disease in both animals and humans . While dog vaccination can interrput transmission in the reservoir, human deaths can also be prevented through prompt administration of post-exposure prophylactic vaccines (PEP) following a bite by a rabid dog [@RabiesVaccinesWHO2018]. However, access to human vaccines are limited in countries where canine rabies is endemic [@sreenivasanOverviewRabiesPostexposure2019; @tarantolaRabiesPostexposureProphylaxis2017; @hampsonEstimatingGlobalBurden2015], and within countries these deaths are often concentrated in rural, underserved communities [@wentworthSocialJusticePerspective2019].

In 2015, a global framework to eliminate deaths due to dog-mediated rabies by 2030 through a combination of PEP provisioning and dog vaccination was established by the World Health Organization (WHO) and partners [@world2018zero]. Furthermore, in 2018, Gavi, the Vaccine Alliance, added human rabies vaccines to their proposed investment portfolio [@GaviVIS]. From 2021, Gavi-eligible countries will be able to apply for support to expand access to these vaccines, with potential to greatly reduce deaths due to rabies. Yet it is unclear how these vaccines should be distributed to prevent deaths most effectively.

A primary challenge is the lack of data on rabies exposures in humans and incidence in animals in most rabies-endemic countries [@scottAddressingDisconnectEstimated2017]. The majority of rabies burden studies at the national level have integrated bite patient data on reported bites into a probability decision tree framework, with the key assumption that overall reported bite incidence (i.e. both bites due to non-rabid and rabid animals) are proportional to rabies incidence (i.e. the more bites reported in a location, the higher the incidence of *rabies* exposures), and that reporting is uniform across space [@hampsonEstimatingGlobalBurden2015; @knobelReevaluatingBurdenRabies2005; @rajeevHealthcareUtilizationProvisioning2018]. If applied subnationally, these assumptions would likely underestimate rabies deaths in places with poor access to PEP and may overestimate rabies deaths in places with better access to PEP.

In Madagascar, the Institut Pasteur de Madagascar (IPM) provides PEP at no-cost to patients at 31 clinics, or anti-rabies medical centers (ARMC) across the country [@rajeevHealthcareUtilizationProvisioning2018]. PEP is not available at any other public clinics or through the private sector. In addition, there is limited control of rabies in dog populations and the disease appears to be endemic throughout the country [@andriamandimbySurveillanceControlRabies2013; @reynesLaboratorySurveillanceRabies2011]. Due to the spatially restricted nature of PEP distribution points and lack of direct costs for PEP, geographic access is likely to be a major driver of disease burden within the country. Previously, we estimated the burden of rabies in Madagascar nationally using data from a single district to extrapolate to the country, and did not account for spatial variation in access [@rajeevHealthcareUtilizationProvisioning2018]. Here, we revise estimates of burden by incorporating the impact of access to PEP at the sub-national level on preventing human rabies deaths and explore the potential impact of expanding provisioning of human rabies vaccines on further reducing these deaths.

# METHODS

## Estimating travel time to closest ARMC {#methods-spatial}
To estimate travel times to the nearest clinic at the scale of administrative units, we used two raster inputs: 1) the friction surface from the Malaria Atlas Project [@weissGlobalMapTravel2018] at an ~1 km^2 scale ([Fig S1.1A](#figS1.1)) and 2) the population estimates which we resampled to the friction surface from the the 2015 UN adjusted population projections from World Pop (@linardPopulationDistributionSettlement2012a, originally at an ~100m^2 resolution, [Fig S1.1B](#figS1.1)).  
Using the approach described in @weissGlobalMapTravel2018, we used GPS points of ARMC to estimate the travel time to the nearest clinic provisioning PEP for the country of Madagascar at an approximately 1 x 1 km scale. We then extracted the mean of travel times for each administrative unit (district commune), as well as the mean of travel times weighted by the population in that grid cell. We also compared these metrics to euclidean distance, but taking the minimum distance from the administrative unit centroid to an ARMC. We used administrative shapefiles from the UN Office for the Coordination of Humanitarian Affairs.

We compared these travel time and distance estimates to driving time data collected by IPM during field missions and patient reported travel times from a subset of patients reporting to the Moramanga ARMC (see Fig S1.2 for raw data).

## Estimating average annual bite incidence {#methods-bitedata}
We used two datasets on bite patients reporting to ARMC:
1) A national database of bite patient forms submitted to IPM from ARMC across the country between 2014 - 2017. These were individual patient data forms that were submitted from clinics as frequently as monthly to annually that included data on the date of reporting and were resolved to the district level (i.e. the district where the patient resided)
2) `r max(mora_bites$nobs, na.rm = TRUE)` months of data (between October 2016 and June 2019) on patients reporting to the Moramanga ARMC resolved to the commune level (the administrative level below the district). We calculated average annual bite incidence to the commune level by taking the monthly average of reported bites and multiplying by 12.

For the national data, some clinics did not report any data, and for others there were substantial periods, months to even a whole year, when no patient forms were submitted. To correct for under-submission of forms, we attempted to identify what proportion of data were missing each year by excluding any periods of time for which there were 15 consecutive days where zero patient records were submitted ([Figure S2.1A](figS2.1). We validated this cut-off of consecutive days by comparing estimated vial demand given the total reported bites (corrected for under-submission) for the years 2014-2016 to actual vials provisioned to clinics over this same period. We did not have the date or month when vials were provisioned, so we compared the totals across this three year period.

Vial demand was simulated under simplified assumptions of PEP administration and adherence[@rajeevHealthcareUtilizationProvisioning2018], assuming that patients reported randomly across the year and then completed doses 3, 7, and 28 days subsequent from the date reported. For this analysis, we look at the range of vial estimates assuming that all patients complete 3 vs. 4 doses total. During this period, the Thai Red Cross Intradermal regimen was used across Madagascar, with 0.2 mL administered per patient at each visit. Vials can be shared within a day between two patients, resulting in 0.1 background wastage per vial shared + any additional wasteage from unused doses discarded at the end of the day. While these simplifying assumptions do not capture variation in adherence, completion, delays, or errors in administration, PEP completion is overall high in Madagascar and in other settings where PEP is provided at no direct costs to patients, so our estimates at 3 doses per patient vs. 4 are likely encompassing much of this variation at the annual level.

We then divided the total number of bites reported to each clinic from each clinic in a given year by the undersubmission rate for eachclinic. Due to the variation in submission rates across time, we chose to take the average of the annual bite incidence estimates aggregated to the district level. The data from the Moramanga district were completely reported across the `r max(mora_bites$nobs, na.rm = TRUE)` month period.

At both the commune and district scale, we assigned clinic catchments by determining which clinic was closest in terms of travel times for the majority of the population within the administrative unit. For the national data, we excluded any districts in a catchment of a clinic which submitted less than 10 forms and any years for which we estimated less than 25% of forms were submitted.

## Modeling reported bites as a function of travel time to closest ARMC {#methods-bitemods}
We modeled the number of reported bites as a function of travel time ($T$) in a poisson regression framework as follows: 
$$ \mu_i = e^{(\beta_{t}T_i \ + \ \beta_0 \ )}P_i$$
$$ y_i = Poisson(\mu_i) $$
where $y_i$ is our data on average number of bites reported to a clinic annually and $\mu_i$ is the expected number of bite patients presenting at the ARMC as a function of travel time to the ARMC ($T_i$) and the human population size ($P_i$) (an offset which scales the predicted number of bites) for a given source location (district or commune). We fit this model to both the national data resolved to the district level and the Moramanga data resolved to the commune level in a Bayesian regression framework via MCMC using the R package ‘rjags’. In order to more directly compare estimates between these two datasets, we also modeled the national data with a commune-level travel time covariate ($T_{j}$) so that:

$$\mu_i = \sum \limits_{j=1}^j e^{(\beta_{t}T_j\ + \ \beta_0 _j)}P_j$$

As travel times are correlated with human population size (Figure S3.1), we also compared how well bites were predicted by human population size alone, and in combination with travel times. For the models with population size, we used a model framework with either population size alone ($\mu_i = e^{(\beta_pP_i + \beta_0 )}$) or population size and travel times ($\mu_i = e^{(\beta_{t}T_i + \beta_pP_i + \beta_0 ))}$) as predictors of bites for the Moramanga data, the national data with covariates at the district level, and the national data with covariates at the commune level.

For the models fitted to the national data, we also attempted to capture variation between clinics by modeling a random catchment effect of the form: $B_{0,k} \sim norm(\mu,\ \sigma)$) where $\mu$ is the mean and $\sigma$ is standard deviation and $B_{0,k}$ is the catchment level intercept). We tested whether the catchment random effect was capturing overdispersion rather than any catchment specific effects by comparing the ratio of the sum of squared residuals to the expected $\chi^2_{n - k}$ distribution given a poisson error structure [@gelman2006data].

We compared alternative model structures for predicting reported bites, specifically by including human population size as a predictor of reported bites (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model), as well as by intercept type ( random = random intercept by catchment, fixed = a single fixed intercept was estimated) for the national data (see main Methods for specific formulations). We used priors centered around zero (*norm(0, 10-6)*) except for the precision parameter for which we used an uniform distribution (*unif(0, 100)*).\

The Moramanga data were not fit to models with a random catchment effect, as these were data from a single catchment. We calculated Deviance Information Criterion (DIC, a metric of model fit to data) for each candidate model as well as potential scale reduction factors (psrf, a metric of model convergence) for each covariate in the model and the multivariate psrf for the whole model (cite Gelman book).

To test our models predictions further, we used parameter estimates from each model to generate predictions for data that were not used to fit the model (i.e. we used the estimates from models fitted to the national data to predict the Moramanga data, and used estimates from models fitted to the Moramanga data to predict the national data). Finally, to check how correcting for incomplete submission of forms affected our modeling results, we fitted these same models to the raw data uncorrected for submission (i.e. assuming that data was completely reported) and with a lower cut-off for identifying periods of undersubmission (7 days vs. 15 days, resulting in higher estimates of undersubmission and accordingly bite incidence).

## Modeling the burden of human rabies deaths sub-nationally {#methods-burden}

We estimate deaths due to rabies as a function of the number of bites predicted by our model and estimates of endemic rabies exposure incidence using an adapted decision tree framework (Fig 2). As with previous decision tree studies [@cleavelandEstimatingHumanRabies2002; @hampsonEstimatingGlobalBurden2015], to model uncertainty in parameter estimates we used triangular distributions for two key parameters: $E_i$, the annual human exposure incidence in a given administrative unit, and $p_rabid$, the proportion of reported bites that are rabies exposures. The range for these distributions was taken from a previous study in Moramanga and also informed by recent methods for estimating underlying rabies incidence in an endemic setting when data is lacking([@whorabiesmodelingconsortiumPotentialImpactImproved2019]).

Specifically: annual rabies incidence of 1% in the dog population in an endemic setting with no mass dog vaccination . To generate $E_i$, the human exposure incidence in a given administrative unit (where $i$ is the index of the commune or district) based on the model, we multiply this baseline incidence by the estimated dog population and the exposure rate per rabid dog ($p_{exp}$). We estimate the dog population given a human:dog ratio (HDR) of 7 to generate our maximum expected incidence and an HDR of 35 for our minimum expected incidence, giving a range of 11.1 ($E_{min}$) - 55.7 ($E_{max}$) exposures per 100,000 persons. As there is little data on dog populations in Madagascar [@andriamandimbySurveillanceControlRabies2013], this range encompasses both a wide range of observed HDRs [@knobelReevaluatingBurdenRabies2005; @sambo2018estimating] and generates similar human exposure incidences as observed previously across Africa [cite here!] . $E_i$ is then drawn from a uniform distribution between the minimum and maximum expected number of human exposures for each location.

```{r fig 1, fig.fullwidth=TRUE}
include_graphics(here("figs/main/M1_sankey.jpeg"))
```
*__Figure 2.__Decision tree for burden estimation. For a given administrative unit $i$, human deaths due to rabies ($D_i$) are calculated from model predicted bites ($B_i$). To get $R_i$, the number of reported bites that were rabies exposures, we multiply $B_i$ by $p_{rabid}$, the proportion of reported bites that are rabies exposures. $p_{rabid}$ is constrained such that reported rabies exposures cannot exceed total rabies exposures ($E_i$) and reporting cannot exceed a maximum ($p_{max}$). $R_i$ is subtracted from $E_i$ to get the number of unreported bites ($U_i$) and then multiplied by the probability of death given a rabies exposure ($p_{infect}$) to get the number of deaths due to rabies ($D_i$). Similarly, deaths averted by PEP are estimated by multiplying $R_i$ by $p_{infect}$, i.e. those who would have died given exposure, but instead received PEP. Both $E_i$ and $p_{rabid}$ are drawn from a triangular distribution; all parameter values and sources are listed in Table 1.*

$B_i$, the number of total reported bites, is the mean predicted number of bites from the models of bite incidence as a function of travel times. To get $R_i$, the number of reported rabies exposures, we multiply $B_i$ by $p_{rabid}$, the proportion of reported bites that are rabies exposures, drawn from a triangular distribution with range between 0.2 - 0.6. So that rabid reported bites cannot exceed the total expected number of human rabies exposures or a maximum reporting (as we know that even with minimal travel times, people may not report for PEP for other reasons), we constrain $p_{rabid}$ as follows:
$$ p_{rabid}=\begin{cases} x, & \text{if}\ \frac{E_i\rho_{max}}{B_i} > x \\ \frac{E_i\rho_{max}}{B_i}, &\text{otherwise} \end{cases}$$
where $\rho_{max}$ is the maximum reporting. We estimate $\rho_{max}$ from the Moramanga ARMC data for the commune of Moramanga Ville, the closest commune to the ARMC (average of 3.12 minutes travel time to the clinic), and where contact tracing data estimated that approximately 2% of rabies exposure go unreported [@rajeevHealthcareUtilizationProvisioning2018]. We assume that all rabies exposed patients who report to an ARMC receive and complete PEP, and PEP is completely effective at preventing death due to rabies. 
We subtract $R_i$ from $E_i$, the number of total human rabies exposures drawn from a triangular distribution with range of 15.6 - 76 exposures per 100,000 persons, to get the number of unreported exposures ($U_i$). Deaths due to rabies in each administrative unit ($D_i$) are then estimated by mutliplying $D_i$ times $p_{death}$.
Deaths averted are calculated similarly as the reported rabies exposures $R_i$ that would have died in the absence of PEP given $p_{death}$ (i.e $A_ip_{death}$). Table 1 lists all parameter values and sources used in the decision tree. \

Table:**Table 1.** Parameters used in the decision tree to estimate human rabies deaths at the administrative level.\

| **Parameter**   | **Value**        | **Description**	                 | **Source**|
|:----------------|:----------------:|:----------------------------------|:---------:|
| $E_i$ |	Triangular(a = 15.6 b = 76, c = 42)| Annual exposures per 100,000 persons	| [@whorabiesmodelingconsortiumPotentialImpactImproved2019; @rajeevHealthcareUtilizationProvisioning2018], see S4|
| $p_{rabid}$	| Triangular(a = 0.2 b = 0.6, c = 0.4) |	Proportion of reported bites that are rabies exposures | [@rajeevHealthcareUtilizationProvisioning2018] |
| $\rho_{max}$ | 0.98	| The maximum reporting possible for any location	Data from Moramanga Ville, Moramanga District | [@rajeevHealthcareUtilizationProvisioning2018] |
| $p_{death}$ |	0.16 | The probability of death given a rabies exposure	| [@changaluchaNeedImproveAccess2018]|

## Estimating the impact of expanding PEP provisioning to additional clinics {#methods-incremental}

```{r ncsbs}
## candidate points
csbs <- read.csv(here("data/raw/csbs.csv"), stringsAsFactors = FALSE)
csbs %>% 
  filter(type == "CSB2", genre_fs != "Priv", type_fs != "Health Post") %>%
  dplyr::select(CTAR = nom_fs, X_COORD = ycoor, Y_COORD = xcoor) -> csbs
```

We developed a framework to rank clinics by how much they could improve access for people currently underserved (> 3 hours away from any clinic in the country) and to estimate the incremental reduction in burden as PEP is expanded to these clinics. We use the approximate locations of all existing public CSB II (centre de sante niveau II, the level below district hospitals, n = `r nrow(csbs)`) as locations of potential ARMC.

Starting with the current locations, we added one clinic at a time, calculating the proportion of people living < 3 hours away from any clinic for the country. We added the clinic which minimized this metric and then repeated the process iteratively, ranking clinics and adding the top clinic sequentially. In this iterative process, we also exclude any clinics that once added only reduced travel times below the 3 hour threshold for less than 0.01% of the population (< `r round(sum(mada_districts$pop)*0.0001, -2)` persons). We also explore the scenario in which all CSB II in the country provision PEP to get an estimate of the maximum impact of expanding PEP provisioning. Resulting travel time estimates and locations of when and where clinics were added are shown in [Supplementary section S5](#S5). Finally, we estimate the incremental reduction in burden as each clinic is added using the burden framework described in the previous section.

We estimate how expansion to additional clinics may impact vaccine demand. Specifically, we use our model-predicted estimates of annual bites at the district and commune level and then aggregate these to the clinic level. As clinics are added, multiple clinics may serve a single district or commune (Fig TO CHECK S5.3B). Therefore, we allocated bites to each clinic according to the proportion of the population in each administrative unit for which it was the closest (Fig CHECK maps). For each clinic, we simulate throughput by randomly assigning patient presentation dates, and then assume perfect compliance (i.e. all patients reporting 3 and 7 days after their first dose) to generate dates of subsequent doses. We use these dates to estimate vial usage given routine vial sharing practices in Madagascar (i.e. each vial is split between two patients, and vials are discarded at the end of each day) and the new abridged intradermal regimen recommended by WHO (2 injection sites on day 0, 3, and 7, [@tarantolaIntradermalRabiesPostexposure2018]). We assume that vial sharing is functioning perfectly, with a background wastage of 0.1 mL per vial, but not accounting for wastage due to other reasons (i.e. wastage due to injection error). We take the mean of 1000 simulations of throughput for each catchment and set of clinics.

## Sensitivity of burden estimates to decision tree parameters {#methods-sensitivity}
To test the effect of our model assumptions on estimates of rabies burden and vial demand, we did a univariate sensitivity analyses of both parameters from the models of bite incidence and models of burden. Additionally, we also examine the systematic variation in human rabies exposure incidence with human population (a potential proxy for changes in the dog population, see section S6) by looking at how estimates of burden change if rabies incidence were to scale positively or negatively with human population size. 
We also generate scaling factors to look at how systematic variation in rabies incidence with human population size could affect our estimates of burden (Fig S6.1). Specifically, if we believe that human to dog ratios are positively correlated with human populations (i.e. if dog ownership/populations are higher in more populated, urban areas), then we might expect higher rabies exposure incidence within our range as population size increases. Alternatively, if we believe that the human to dog ratio is inversely correlated with human population size (i.e. if dog ownership is more common in less populated, rural areas), we would expect rabies exposure incidence within our range to scale negatively with human population size.

## Data and code 
All analyses were done in `r sessionInfo()$R.version$version.string` [@RLang2019] and using the packages listed in the [supplementary references]{#srefs}. All processed data, code, and outputs are archived at this zenodo repository and maintained at this github repo. The raw bite patient data at the national level is maintained by IPM. This database was last queried on September 19, 2019 for these analyses. Anonymized raw data from this database are available from IPM following insitutional data transfer protocols. Anonymized raw data collected from the Moramanga District were retrieved from the Wise Monkey Portal (http://www.wisemonkeyfoundation.org/) on the same date and are shared in the archived repository.

## Ethics statement
Data collection from the Moramanga District was approved by the Princeton University IRB (# 7801) and the Madagascar Ministry of Public Health Ethics Committee (# 105-MSANP/CE). Oral informed consent was obtained from all interviewed participants. Data collected from bite patients at the national level are maintained jointly by the Ministry of Health and IPM as a routine part of PEP provisioning.

# RESULTS

## Estimating travel times to the closest ARMC {#results-ttimes}
```{r ttime stats}
ttime_stats <- read.csv(here("output/stats/prop_pop_ttimes.csv"))
# get prop of pop > 3 hrs away at baseline
ttime_stats %>%
  filter(scenario == "baseline", !(cut_times %in% c("0-1", "1-2", "2-3"))) %$%
  sum(prop_pop) -> prop_over3_base
ttime_stats %>%
  filter(scenario == "baseline", !(cut_times %in% c("0-1", "1-2", "2-3", "4-5", "5-6"))) %$%
  sum(prop_pop) -> prop_over6_base
```

Estimates of travel times to the closest ARMC vary greatly across the country (Fig 3A), with approximately `r round(prop_over3_base*100, 0)`% of the population living greater than 3 hrs away from a clinic (Fig 3B). We found that travel times estimated from the friction surface generally underestimate true driving times and patient reported travel times to the Moramanga ARMC (Fig 3C), and that patient reported travel times were significantly more variable than estimates from the friction surface (Fig S1.2). The friction surface assumes that given multiple types of transport in a cell, the fastest mode of transport takes precedence (i.e. if there is a road, then people will travel at vehicle rather than walking speeds). However, patients reporting to the Moramanga ARMC report using mutliple modes of transport, with some cases of individuals walking multiple days to reach a clinic (Fig S1.3). Despite these discrepancies, travel times correlate with the reported and groundtruthed driving times (Fig 3C, Fig S1.4). Travel times weighted by population at the grid cell level were a better predicter than unweighted travel times or distance (Table S1.1). Therefore we use this metric as a proxy for access at the commune/district level in subsequent analyses.

```{r fig M2, fig.fullwidth=TRUE}
include_graphics(here("figs/main/M2_ttimes.jpeg"))
```
*__Figure 3.A)__ Travel times to the closest ARMC estimated at an ~ 1km^2 scale for Madagascar. __B)__ Distribution of the population across travel times. __C)__Correlation between patient reported travel times to the Moramanga ARMC and travel times weighted by population (mean at the commune level) and between reported driving times between GPS points and travel time estimates at the grid cell level. For the commune means, the vertical point lineranges show the 95% quantiles for reported travel times, and the point size shows the number of observations for each commune. The best fit lines (pink and purple lines) and 95% confidence interval (shading) from a linear model where observed travel times ~ estimated travel times for each data source are also shown. The dashed black line is the 1:1 line.*

## Estimating average annual bite incidence at the district and commune level {#results-biteinc}
We assigned each district or commune to a catchment based on which clinic was the closest for the majority of the population. While there are discrepancies between commune and district catchment assignments (Fig S2.1A), > 75% of the population in a given district or commune were closest to a single clinic (Fig S2.B). In addition, for the national data we found that for most districts, the majority of bites were reported to the closest clinic as estimated by our travel time metric (Fig 1A), and the majority of bites reported to a clinic came from within the catchment assigned by the travel time metric (Fig S2.1C).

```{r fig M3 catchmaps}
include_graphics(here("figs/main/M3_bitemaps.jpeg"))
```
*__Figure 3.A)__ The network of patient presentations at the district level for the national data: circles with black outline represent the total number of patients reporting to each ARMC for which we have data. Color corresponds to the clinic catchment. Circles with white outline are the total number of bites reported for that district. Lines show which ARMC those patients reported to, with the line width proportional to number of patients from that district reporting to the ARMC; flows of less than 5 patients were excluded. Districts in catchments excluded due to lack of forms submitted by the clinic are colored in grey. Out-of-catchment reporting is indicated where points and line colors are mismatched. __B)__ The network of patient presentations at the commune level for the Moramanga data with inset of Madagascar showing the location of the area plotted. The black outline shows the district of Moramanga, and the red polygons shows all the communes included in the Moramanga ARMC catchment. Circle with black outline is the location of the Moramanga ARMC with the size of points proportional to the total number of bites reported to the clinic, as per A.*

Clinic submission of forms was highly variable over the period covered by the dataset (Figure S2.1). We found that estimates of vial demand based on uncorrected bite patient numbers are generally lower than the true number of vials provisioned for those clinics with substantial undersubmission of forms ([Figure S2.1B](#figS2.1)), and correcting patient numbers by under-submission of forms based on the 15 day consecutive day cut-off results in estimates of vial demand closer to the observed vials provisioned ([Figure S2.2A and B](#figS2.2)). We excluded any clinics which submitted less than 10 forms (excluded `r nrow(ctar_metadata) - max(district_bites$catch)` catchments, in grey in Fig 3A). After additionally excluding any year with less than 25% of forms submitted, our final dataset consisted of estimates of average bite incidence for `r nrow(district_bites)` out for `r length(mada_districts)` districts total (Fig 4A), and `r nrow(mora_bites)` communes within the catchment of the Moramanga District (Fig 4B).

```{r fig M4 biteinc}
include_graphics(here("figs/main/M4_biteinc.jpeg"))
```

*__Figure 4.__ The estimated average annual bite incidence from the national and Moramanga data plotted at the __A)__ district scale (both datasets) and at the __B)__ commune scale (Moramanga dataset). Colors correspond to the clinic catchment, shape indicates the dataset, and the size of the point indicates the number of observations (i.e. the number of years for which data was available for the national data; note for Moramanga `r max(mora_bites$nobs, na.rm = TRUE)` months of data were used). The point lines indicate the range of estimated bite incidence for each district.*

```{r}
outliers <- read.csv(here("output/stats/outliers.csv"))
```

Bite incidence estimates generally decreased with weighted travel times at both scales, although there was considerable variation between catchments for the magnitude of this relationship. For the national data, there were also two outliers, Toamasina II (the sub-urban commune surrounding the city of Toamasina) and Soanierana Ivongo, which both had higher bite incidence when compared to other districts with similar travel times. While the estimates from the Moramanga data when aggregated to the district fell within the ranges observed from the National dataset, at the commune scale they showed more variation, with higher reported incidence at low travel times at the commune level compared to the district.

## Modeling reported bite incidence as a function of travel times {#results-bitemods}
All models and posterior estimates of parameters converged ({inline stat for max psrf & mpsrf}). Population size was the poorest predictor of bites (Fig S3.2, Table S3.1) and adding population size to the model did not improve the fit to the national level data (Figure S3.2), although these models did have lower DIC values for the Moramanga data (Table S3.1). However, when used to predict data that were not used to fit the models (i.e. model estimates from fitting to the Moramanga data were used to predict the national data and vice versa), only models with incidence offset by population size (flatPop) rather than scaling with population size (addPop) generated reasonable predictions to the data (Figure S3.3).\

All of the candidate models which included travel times produced similar predictions when compared to the observed data (Fig S3.2). Population size alone was the poorest fit to the data as estimated by DIC (Table S3.1), and models with population size as an additional covariate did not generate realistic predictions to the observed data or when used to predict out of fit (Figs S3.2 and S3.3). For the national data, the models with travel times alone as a covariate and a random catchment generated predictions closest to the observed data (Fig S3.2). These predictions were robust to how we corrected for incomplete submission of data by clinics (Fig S3.5). The models without the random effect were significantly overdispersed, but the catchment random effect still captured additional variation beyond this overdispersion (Fig S3.4). The parameter estimates from the different models were similar, with the model fit to the Moramanga data showing the highest impact of travel times and the district model the lowest (Fig 5B-D), however the increased bite incidence observed at low travel times in the Moramanga data was not captured by the commune model (Fig 5A). To capture the range of estimates of travel time impact, we compared the district model fitted to the national data with the commune scale model fitted to the Moramanga data in subsequent analyses.

```{r fig 5, fig.fullwidth=TRUE}
include_graphics(here("figs/main/M5_mods.jpeg"))
```
_**Figure 5.** Travel times as a predictor of reported bite incidence per 100,000 persons. A) The estimated relationship between travel time in hours (x-axis) and mean annual bite incidence per 100,000 persons (y-axis). The lines are the estimates using the posterior estimates of the parameters for three candidate models: model with travel times at the 1) commune and 2) district level fitted to the national data with a random intercept by catchment, and 3) travel times at the commune level fitted to the Moramanga data with a fixed intercept and adjusted for overdispersion. The envelopes shows the 95% prediction interval. The points show the data used to fit the models (National data (triangles) at the district level used to fit the District and Commune models, and Moramanga data (circles) at the commune level used to fit the Moramanga model). B - D) The posterior distribution of parameters from the respective models for the B) model intercept, C) travel time effect, and D) the catchment random effect (for national data only)._

## Estimating the burden of human rabies deaths subnationally {#results-burden}

```{r burden results}
natl_burden <- read.csv(here("output/preds/natl_burden.csv"))
```

Overall, we estimate between `r round(sort(natl_burden$deaths_mean)[1])` - `r round(sort(natl_burden$deaths_mean)[nrow(natl_burden)])` average annual deaths in Madagascar due to rabies depending on the model used (Table 2). We also estimate that under the current system of 31 ARMCs in Madagascar, approximately `r round(sort(natl_burden$averted_mean)[1])` - `r round(sort(natl_burden$averted_mean)[nrow(natl_burden)])` deaths due to rabies are prevented through provisioning of PEP each year. In general, the incidence of rabies deaths increases with travel times to clinics, and there is significant variation sub-nationally when burden is modeled at the district level, with even stronger patterns of increase in burden with travel times when modeling burden at the commune level with estimates from the Moramanga data (Figure 4). The estimates of burden from the model fitted to the Moramanga data result in estimates of burden that are lower at the national level, but higher relatively in places with high travel times {fix!}. However, there is considerable uncertainty in these estimates, in particular given the district model of bite incidence.

*__Table 2.__ Model predictions of average annual reported bite incidence, total deaths, and total deaths averted at the national level for the two models (commune level and district level models with a random catchment effect); 95% prediction interval in parentheses.*

```{r Table 2}
natl_burden %>%
  mutate_at(vars(starts_with("bites")), function(x, p) x/p*1e5, p = .$pop) %>%
  mutate_at(vars(bites_mean:averted_lower), round) %>%
  mutate(scale = case_when(scale == "Commune" ~ "Commune (Moramanga data)",
                           scale == "District" ~ "District (National data)"), 
         bites_100k = glue("{bites_mean} ({bites_lower} - {bites_upper})"),
         deaths = glue("{deaths_mean} ({deaths_lower} - {deaths_upper})"),
         averted = glue("{averted_mean} ({averted_lower} - {averted_upper})")) %>%
  select(scale, bites_100k, deaths, averted) -> table1

kable(table1, digits = -0, col.names = c("Model", "Reported bite incidence per 100k", "Burden of deaths", 
                                        "Deaths averted"))           
```

```{r fig 4, fig.fullwidth=TRUE}
include_graphics(here("figs/main/M6_burden_base.jpeg"))
```
*__Figure 4.__ Spatial variation in predicted incidence of human rabies deaths per 100, 000 persons. __A)__ for each district (y-axis) in Madagascar, where the diamonds show the predicted incidence for the district model fit to the national data and the squares show the incidence for the commune model fit to the Moramanga data for all communes in a district. Points are colored by and districts are ordered by travel times. The vertical lines show the average national incidence of human rabies deaths for the commune (grey) and district (black) models. Incidence mapped to the __B)__ commune level and __C)__ district level from the respective models; grey X’s show locations of current CTAR.*

## The impact of expanding PEP access to additional clinics {#results-incremental}
```{r}
natl_preds_all <- read.csv(here("output/stats/natl_preds_all.csv"))
natl_preds_all %>%
  group_by(scale) %>%
  arrange(scenario) %>%
  mutate_at(vars(-c(group_cols(), scenario)), function(x) x/x[1]) -> props
trends <- read.csv(here("output/stats/trend_preds.csv"))

```

We found that expanding PEP to additional clinics reduces rabies deaths, with strategic expansion of access to an additional 250 clinics reducing burden by almost `r round(min(1 - props$deaths_mean[props$scenario == 250])*100, 2)` - `r round(max(1 - props$deaths_mean[props$scenario == 250])*100, 2)`% (Fig M7.A). However, benefits accrue less rapidly as ARMC are added: with approximately 700 clinics, we see a further reduction between `r round(min(1 - props$deaths_mean[props$scenario == 700])*100, 2)` - `r round(max(1 - props$deaths_mean[props$scenario == 700])*100, 2)`, while even when all CSB IIs in the country provision PEP, the most deaths are reduced is between `r round(min(1 - props$deaths_mean[props$scenario == max(props$scenario)])*100, 2)` - `r round(max(1 - props$deaths_mean[props$scenario == max(props$scenario)])*100, 2)`%. This saturation is a result of the landscape of clinics across the country, as most CSBs are in areas with already low travel times (Fig S5.X). Therefore, as clinics are added, reported bite incidence saturates (Fig S5.Z), and instead there are shifts in which clinics bites are reported to (Fig S5.Y). Even at this maximal PEP access, our model still predicts `r round(min(1 - natl_preds_all$deaths_mean[natl_preds_all$scenario == max(natl_preds_all$scenario)]), 2)` - `r round(max(1 - natl_preds_all$deaths_mean[natl_preds_all$scenario == max(natl_preds_all$scenario)]), 2)` rabies deaths per annum, and average reporting of rabid exposures still remains between `r round(min(1 - trends$reporting_mean[trends$scenario == max(trends$scenario)])*100, 2)` - `r round(max(1 - trends$deaths_mean[trends$scenario == max(trends$scenario)])*100, 2)`% (Fig S5.N).

Vial demand also increases, and more vials are needed per averted death as PEP access is expanded. Our model predictes an increase from `r round(min(1 - natl_preds_all$vials_mean[natl_preds_all$scenario == 0]), 2)` - `r round(max(1 - natl_preds_all$vials_mean[natl_preds_all$scenario == 0]), 2)` vials under current clinic provisioning and with the abridged Thai Red cross regimen (i.e. visits on days 0, 3, 7), to ~`r round(min(1 - natl_preds_all$vials_mean[natl_preds_all$scenario == 250]), 2)` - `r round(max(1 - natl_preds_all$vials_mean[natl_preds_all$scenario == 250]), 2)` {inline} vials with 250 clinics, and approximately `r round(min(1 - natl_preds_all$vials_mean[natl_preds_all$scenario == max(natl_preds_all$scenario)]), 2)` - `r round(max(1 - natl_preds_all$vials_mean[natl_preds_all$scenario == max(natl_preds_all$scenario)]), 2)` vials if all CSB IIs were to provision PEP. Most bites become reported within a commune--why does throughput decrease!

Similar to our baseline estimates of burden, there was significant uncertainty in these estimates, in particular for the district model fit to the national data.

<!-- TO DO: State what happens to throughput + vial sensitivity! -->

```{r fig 5, fig.fullwidth=TRUE}
include_graphics(here("figs/main/M7_addARMC.jpeg"))
```

*__Figure 5. A)__ Decrease in deaths due to rabies, __B)__ increase in total # of vials as additional ARMC are added at the national level , and __C)__ increase in vials needed per death averted based on the two models of reported bites. Lines are the mean of 1000 simulations with envelopes representing 95% prediction intervals. The “...” on the axes indicates a break in the axes, and the last point is the burden and vial predictions when all CSBS in the country provision PEP (maximum access).*

## Sensitivity of estimates to bite incidence and decision tree model parameters {#results-sensitivity}

Overall, our sensitivity analyses show that while there is significant uncertainty in the absolute numbers of deaths due to rabies, patterns with increasing risk of death with reduced access and the potential for expanding access to PEP to reduce these deaths appear to be robust (Figs S6.2, S6.3). In general, for a given incidence level the higher our assumption of the proportion of reported bites which are rabid, the lower the estimated burden. The lower our estimate of maximum reporting, the higher overall burden of rabies, and as this parameter decreases, the relationship between travel times and rabies deaths are dampened. Scaling with population size modulates this impact of travel times on deaths, with positive scaling of rabies incidence with population size (i.e. more rabies in more populated places) dampening and negative scaling enhancing the relationship between access and deaths. Finally, raw numbers of deaths greatly depend on what we assume the rabies exposure incidence is in the population.

To quantify which parameters contribute to the uncertainty observed in our estimates, we performed a univariate sensitivity analyses, varying parameters in both the model of bite incidence (taking the lower and upper 95% confidence intervals of the posterior estimates, Table 1) and the decision tree model (fixing parameters at the ranges used in the parameter distributions, Table 2). These analyses show that assumptions of the underlying rabies exposure incidence ($E_i$) contribute the greatest to uncertainty in estimates of baseline burden (Fig S6.Y). For the impacts of expanded access, what we assume is the maximum reporting of rabies exposures given minimal travel times ($rho_max$) is an important factor for the commune model (Fig S6.X), because as clinics are added, more locations are shifted to this threshold (Fig S5.X). For the district model, $p_{rabid}$ and $E_i$ are also important in determining the magnitude of reduction of burden. Despite the large absolute ranges of parameter estimates, the broad patterns of higher risk with increased travel times are robust for both our estimates of burden and the impacts of expanding access to PEP, even given systematic bias in our assumptions of incidence (Fig S6.N).

Uncertainty in bite model parameters contribute less to variation of estimates of burden. For the estimates of vial demand, uncertainty around the intercept of the model (i.e. the baseline reported bite incidence) has the greatest impact, rather than the travel time effect or the catchment effect (Fig S6.4).

```{r se}
se_pars <- read.csv(here("output/sensitivity/se_pars.csv"))

se_pars %>%
  pivot_longer(exp_min:beta_0) %>%
  mutate(value = case_when(name %in% c("exp_min", "exp_max") ~ round(value*1e5, 2),
                           !(name %in% c("exp_min", "exp_max")) ~ round(value, 2))) %>%
  group_by(scale, par = name) %>%
  summarize(upper = max(value), lower = min(value)) %>%
  mutate(range = glue("{lower} \U2012 {upper}")) -> se_table_all

se_table_all %>%
  filter(grepl("beta|sigma", par)) %>%
  pivot_wider(par, values_from = range, names_from = scale) %>%
  as.data.table(.) -> bites_se
se_table_all %>%
  filter(!grepl("beta|sigma|min", par), scale != "Commune") %>%
  as.data.table(.) -> dtree_se
```

Table:**Table 3.** Range of parameters used in models of bite incidence for the univariate sensitivity analyses.\

| **Parameter**   | **Description**        | **Commune model (Moramanga data)** | **District model (National data)** | 
|:----------------|:----------------------:|:----------------------------------:|:----------------------------------:|
| $\beta_t$ | Travel time effect	| `r bites_se$Commune[par == "beta_t"]` | `r bites_se$Commune[par == "beta_t"]`
| $\beta_0$ | Model intercept	| `r bites_se$Commune[par == "beta_0"]` | `r bites_se$Commune[par == "beta_0"]`
| $\sigma_0$ | Standard deviation of intercept by catchment	| -- | `r bites_se$Commune[par == "sigma_0"]`

Table:**Table 4.** Range of parameters used in decision tree model for the univariate sensitivity analyses.\

| **Parameter**   | **Description**        | **Range**	                 | **Notes**|
|:----------------|:----------------------:|:----------------------------|:---------:|
| $E_i$ |	Annual exposures per 100,000 persons| `r dtree_se$range[par == "exp_max"]`  	
| $p_{rabid}$	| Proportion of reported bites that are rabies exposures | `r dtree_se$range[par == "p_rab_max"]` 	
| $\rho_{max}$ | The maximum reporting possible for any location | `r dtree_se$range[par == "rho_max"]`	 
| $p_{death}$ | The probability of death given a rabies exposure^[1] | `r dtree_se$range[par == "p_death"]`	
[1] 95% CI of estimate from @changaluchaNeedImproveAccess2018|

# DISCUSSION  
```{r}
natl_preds_all %>%
  filter(scenario == 0) %>%
  mutate(prop_averted = averted_mean/(deaths_mean + averted_mean)) %>%
  select(prop_averted) -> prop_averted
```

Notes from Claire's MSc
22 ARMC ask for fees (consultation fees to fees for guaze etc for the form. ranging from 2000 Ar - 10000 Ar in some cases, 0.50 - 3,00 USD)
60% would vaccinate cat I  exposures
3 ARMC were out of stock 
282000 Euros a year on 23800 + 9500 vacc vials, 2000 RIG vials.

## Main findings
Travel time to the closest ARMC was a strong predictor of reported bite incidence in Madagascar. Using an adapted decision tree framework, we find that the burden of rabies is likely to be concentrated in areas with low access to PEP. Even given only 31 clinics, we estimate that current provisioning averts `r glue("{trans_perc(min(prop_averted))} - {trans_perc(max(prop_averted))}%")` of deaths that would occur due to rabies in the absence of PEP. We estimate that expanding PEP access has the potential to greatly reduce deaths due to rabies, with expansion to an additional 200 {inline} clinics potentially reducing burden by `r glue("{trans_perc(min(1 - props$deaths_mean[props$scenario == 250]))} - {trans_perc(max(1 - props$deaths_mean[props$scenario == 250]))}%")`. However, our model predicts that geographic expansion of PEP access alone is unlikely to eliminate deaths due to rabies, and that even given expansion of PEP (i.e. to all major health centers (n = 1676) in the country), between `r glue("{round(min(natl_preds_all$deaths_mean[natl_preds_all$scenario == max(natl_preds_all$scenario)]))} - {round(max(natl_preds_all$deaths_mean[natl_preds_all$scenario == max(natl_preds_all$scenario)]))}")` deaths may still occur nationally, due to the distribution of clinic locations across the country. After a certain number of clinics are added, rather than large increases in reported incidence, we see shifts in the clinic locations where bites are reported to. Vial demand reflects these patterns as well, with more vials needed per death averted as clinics are added, due to the demand for PEP from patients not genuinely exposed to a rabid animal.

## Limitations
The travel time estimates from the MAP friction surface underestimate true travel times reported by patients. This likely has to do with the quality of data on transport speeds in Madagascar, which is taken from the Open Street Map user community, or in it's absence, from country cluster data [@weissGlobalMapTravel2018]. In Madagascar, the presence of paved roads does not necessarily reflect their quality or the modes of transport used to travel these roads. Patients reporting to the Moramanga ARMC reported using a variety of transport methods and had highly variable travel times even within a single commune. Future studies should aim to improve estimates of travel times with country specific data on road speeds and travel conditions [@rayAccessModComputingGeographic2008] to better estimate spatial inequities in access, or by collecting data on patient reported travel times. While patient reported travel times may not be precise, they will likely be more reflective of the range of travel realities that people experience.

The national data we used were underreported and incomplete, but we attempt to correct for these issues (quantifying under-submission of forms and excluding districts from catchments with less than 25% of data reported). Improving data submission and quality from clinics across the country will be important to better understand patterns of health seeking behavior and access, but also to monitor how well clinics are functioning and to develop better provisioning systems. In addition, we found that the scale of data used to fit these models is likely an important factor in estimating the impact of spatial access. The commune level data from the Moramanga District showed a steeper decline in reported bite incidence with travel times and higher estimates of bite incidence at lower travel times, resulting in lower burden estimates overall, but a greater concentration of burden in areas with high travel times. However, we lacked data from other catchments at the commune scale, and the commune model we developed to fit to the district scale data was not able to capture these same patterns from the national data. The district model fitted to the national data shows the more conservative estimate of the impact of current and expanded PEP access on human rabies deaths.

We assume that spatial access to PEP is the primary driver of health-seeking behavior, however many other factors may drive an individual’s decision to seek PEP, such as socioeconomic and educational status and awareness about the danger of rabies [@samboKnowledgeAttitudesPractices2014; @barbosacostaInfluencePovertyRabies2018; @etheartEffectCounsellingHealthcareseeking2017; @hampsonRabiesExposuresPostExposure2008]. In the Madagascar context where PEP is provided free of charge to patients, the main cost of seeking PEP is for transport and associated time lost for travel. In addition, more remote communities tend to be of lower socioecomonic and educational status [@peters2008poverty], so our travel time metric may act as a proxy for these set of correlated variables that impact an individual’s access to PEP. We included a catchment level random intercept to account for spatial differences in reported bite incidence not due to travel times. This random component of the model improved fit to the data, suggesting that variation at the clinic level (i.e. differences in clinic functioning such as vaccine availability) and at the catchment level (i.e. differences in regional bite incidence, awareness, outreach, etc.) are likely important drivers of health-seeking behavior and overall vaccine demand. A recent knowledge-attitudes-practices (KAP) study of both community members in the Moramanga District and ARMC staff across the country found that both socioeconomic status and living in an urban area were predictive of knowledge scores [cite Leblanc 2019]. Outlier discussion?

Our decision tree framework makes strong assumptions about baseline rabies exposure incidence, based on a previous study from the Moramanga District [@rajeevHealthcareUtilizationProvisioning2018] and a global study of rabies burden[@whorabiesmodelingconsortiumPotentialImpactImproved2019]. As incidence of rabies exposures varies significantly over time and space [@hampsonSynchronousCyclesDomestic2007; @hampsonTransmissionDynamicsProspects2009], we incorporated uncertainty into these estimates to generate a mean estimate of burden. Even when we assume the extremes of these ranges, or that rabies exposure incidence varies systematically with human populations, we find that the qualitative effects of travel times on risk of rabies deaths remains the same. While our sensitivity analyses show that the relative patterns of burden are robust, the magnitude of deaths is most sensitive to our assumption of rabies exposure incidence. Thus, our estimates are more representative of patterns of risk than quantitative differences in burden, and spatial estimates of rabies exposure incidence would be necessary to generate more precise subnational estimates of burden.

We also make several simplifying assumptions regarding PEP compliance and efficacy. First, we assume perfect compliance of patients and efficacy of PEP. Although delays in reporting and poor adherence to PEP contribute to risk of death, PEP compliance is generally high in Madagascar [@rajeevHealthcareUtilizationProvisioning2018], and these factors are unlikely to be major drivers of burden [@changaluchaNeedImproveAccess2018]. We also assume that all ARMC reliably provision PEP, however Leblanc 2019 found that at the time of the KAP survey, three clinics were out of vaccine stock which likely contributes to burden. We do not account for differential risk for severely exposed patients (i.e. Category III exposures) that do not receive rabies Immunoglobulin (RIG). RIG is only available at IPM in Antananarivo, and while exposed persons may be referred to seek RIG, the majority of Category III exposures that seek care are unlikely to have received RIG. Unfortunately, we were not able to link patient records between peripheral ARMC and IPM to directly estimate this. Recent studies have shown that even in the absence of RIG, PEP is highly effective at preventing rabies deaths, and only adds a marginal benefit in terms of reducing burden but at very high cost. Therefore, it is both unlikely that availability of RIG contributes significantly to rabies burden in Madagascar, or that expansion of RIG provisioning would be a feasible option given the costs.

Ideally, many of these assumptions could either be validated by or replaced with improved data on canine rabies incidence, health-seeking behavior, and clinic functioning and provisioning of PEP. However, even with these data sources, it will be challenging to incorporate the complex interactions between socioeconomic factors, access to and quality of care, and human behavior to predict rabies burden. We detail case studies in Box 1 that highlight some of the challenges in capturing variable healthcare seeking behavior, particularly in the context of PEP provisioning, and the policy implications of these exceptions.

## Box 1: case studies of the challenges of predicting health seeking behavior

### Anosibe A'nala
In the district of Anosibe A'nala (population of ~ xx people) south of the Moramanga District, travel times to the nearest ARMC (the Moramanga ARMC) are high and highly variable, often exceeding 24 hours and consisting of days of travel on foot. Despite these barriers, we see a high incidence of reported exposures (~~ xx per 100,000 persons). We interviewed xx% of these patients when they reported to the clinic, and about xx% {inline} of persons interviewed were loosely defined Category I exposures, members who were in close proximity or had touched a person that had died from suspected rabies. Approximately xx% of non Category exposures were classified as a probable rabies case based on clinical criteria. No cases were confirmed diagnostically however. The data from Anosibe A'nala points to three key insights: 
1) This remote community does suffer from rabies exposure and is likely experiencing a high burden of death that we are only seeing the tip of the iceberg for. 
2) When people are aware of the risk of death due to rabies, they are willing to travel a long time (in many cases walking 3 days to attend the clinic) to seek the vaccine. In addition, surveillance responses, i.e. when there is a suspected cases, there is often referral of these people from far away, can work and if these investigations are triggered on exposure rather than death then they could greatly help to prevent deaths. There is great potential for community outreach to go far. 
3) The travel times estimated from the friction surface are much lower than those observed. Because while there is a road, road speeds are often < 10km per hour and people hitch rides with transport trucks. This is also highly seasonal, during the rainy season it's almost impossible to travel other than by foot. And most people use multiple modes of transport depending on cost and availability (i.e. boat, on foot, by camion). So these metrics may miss key areas that are underserved and warrants improvement.

### Antananarivo 
A middle aged taxi driver died of rabies in suburban Antananarivo in 2016. Bitten by an unknown dog. Promptly reported to a clinic where he was reffered to IPM. His daughter, a prospective medical student, and other family members urged him to seek the vaccine but he did not in the end seek care. He developed symptoms 2 weeks later and died within a few days. The case was confirmed by the Naional Rabies Reference Laboratory at IPM. This case points to the other side of the health-seeking behavior coin from AAA. This man had the resources, was aware of the risk, and was reffered by the hospital (and sought care immediately). But ultimately, it's difficult to predict human behavior. Ultimately it will be difficult to acheive 100% coverage of this intervention even in the best of settings. Why we don't see people dying in other places is because of the combo of dog vacc + PEP vacc.

## Broader context

Our work is in line with previous studies demonstrating the impacts of geographic access to care on health outcomes [@aleganaTreatmentseekingBehaviourLow2017; @nikolayEvaluatingHospitalBasedSurveillance2017; @macphersonDisparitiesAccessDiagnosis2019; @polettiHiddenBurdenMeasles2018; @Hegde2019dt; @metcalfTransportNetworksInequities2015]. While many of these studies directly estimated reporting using community based and cross-sectional survey data [@aleganaTreatmentseekingBehaviourLow2017; @nikolayEvaluatingHospitalBasedSurveillance2017], we use model derived estimates of reporting and assumptions of baseline disease incidence to estimate burden, similar to methods developed for other pathogens to estimate burden in data sparse settings [@macphersonDisparitiesAccessDiagnosis2019; @polettiHiddenBurdenMeasles2018; @Hegde2019dt]. For canine rabies, estimating burden of disease directly is challenging [@scottAddressingDisconnectEstimated2017; @taylorDifficultiesEstimatingHuman2017] and direct empirical estimates of rabies burden have previously relied on exhaustive contact tracing studies [@beyeneEstimatingBurdenRabies2018; @hampsonRabiesExposuresPostExposure2008] or large-scale mortality surveys [@sudarshanAssessingBurdenHuman2007]. Recent studies have estimated access to health-seeking behavior and PEP completion and adherence, but not directly linked these estimates to burden [@tarantolaRabiesPostexposureProphylaxis2017; @zaidiGeographicVariationAccess2013]. Our framework for incorporating estimates of access into previously developed burden estimation methods (i.e. decision trees) could help to improve and refine existing burden estimates and explore impacts of expanding access, particularly at the sub-national level.

The observed range of reported bite incidence from our data are in line with estimates of reported bite incidence across Africa, and largely is in line with previous model estimates of bite incidence based on national statistics (i.e. human development index & proportion of the population that is urban, see @@whorabiesmodelingconsortiumPotentialImpactImproved2019 Supplementary Materials for both a summary of data sources and model estimates). Ultimately, reported bite incidence is driven by many factors unrelated to rabies including the underlying size of dog populations, barriers to access to PEP (i.e. both financial costs and geographic access), and awareness of rabies risk. In our study, we do not attempt to quantify the total non-rabies related bite incidence, but use a range of values for the proportion of reported bites which we think are due to rabies. The approach we used here is likely less useful in settings where rabies exposures only make up a small fraction of patients reporting for PEP [@rysavaPathRabiesElimination2018; @wallaceEstablishmentCanineRabies2015], i.e. in settings where barriers to PEP are low (i.e low or no direct costs, low travel times, high socioeconomic status of the population).

While our incidence of human rabies deaths are based on strong assumptions of rabies incidence, they generally fall within the range estimated for a wide range of settings from both empirical and modeling studies (1 - 7 subnationally, on average 2.7 - 3.1, see [@beyeneEstimatingBurdenRabies2018; @hampsonRabiesExposuresPostExposure2008; @sudarshanAssessingBurdenHuman2007; @tenzinDogBitesHumans2011; @lyRabiesSituationCambodia2009, @whorabiesmodelingconsortiumPotentialImpactImproved2019]. Our estimates of rabies burden fell within the higher end of the range of previous estimates of burden for Madagascar [@rajeevHealthcareUtilizationProvisioning2018]. While we assumed a lower range of incidence of underlying rabies exposure, accounting for spatial variation in access and in reported bite incidence across the country resulted in predictions of increased burden, mirrored by a decrease in deaths averted under baseline provisioning. However, our results covered wide 95% prediction intervals, and improved clinic level data, as well as surveillance data on rabies exposures will be necessary to further refine these absolute estimates.

One approach to obtaining this data is integrated bite case management and clinic-based surveillance, methods that relies on follow-up of bite patients reporting to determine rabies exposure status, recommend treatment, and identify other potential persons exposed [@lechenneImportanceParticipatoryIntegrated2017; @undurragaCostEffectivenessEvaluationNovel2017; @rajeevHealthcareUtilizationProvisioning2018]. If these methods were implemented, ARMC could act as sentinels for rabies surveillance across the country, with bite patients reporting to clinics trigger investigations of suspect rabies cases. Ultimately, these methods rely on exposed persons to report to clinics, and as such, expanding access to PEP could strengthen clinic based surveillance efforts.

Our estimates of vial demand depend on countrywide adoption of the new abridged intradermal regimen recommended by WHO [@tarantolaIntradermalRabiesPostexposure2018]. While this new recommendation has been adopted by the Ministry of Health, it will be important to train clinicians and staff, as in 2019 only 15% of staff interviewed were aware of this regimen [Leblanc 2019]. In addition, the majority of staff were also not aware of the WHO classifications of exposure categories, and vaccination of Category I exposures, i.e. those not recommended PEP, is a common practice. In the Moramanga district, up to 20% of vial demand was due to these exposures [@rajeevHealthcareUtilizationProvisioning2018]. More judicious PEP provisioning could reduce vial demand. Current practices of provisioning PEP reactively to Category I exposures. Based on data from the Moramanga District, vial demand could be up to 20% lower if these practices were stopped.

We predict that as clinics are expanded, overall throughput (i.e. the number of patients reporting daily to a clinic) will decrease. This may make provisioning PEP more challenging, as vial demand may become less predictable annually, and result in stockouts or vaccine wastage. Establishing decentralized provisioning mechanisms, for example through combining provisioning of PEP with routine childhood vaccines, or novel methods for vaccine delivery (i.e. drone delivery of emergency medicine, [@boeras2019use]) may be potential ways to mitigate these challenges. Previous to 2006 when the nerve tissue vaccine was still in use, access to PEP in Madagascar was more widespread, with clinics able to request vaccine from IPM upon demand. However, once the more expensive Vero cell vaccine was implemented, vaccine provisioning to all clinics became too costly, resulting in provisioning to a subset of clinics [@andriamandimbySurveillanceControlRabies2013]. This suggests that wider spread vaccine provisioning is feasible if cost barriers are removed.

A Gavi investment can remove this cost barrier. In most rabies endemic countries, access to PEP is geographically limited and/or costly to patients [@changaluchaNeedImproveAccess2018; @liDescriptiveAssessmentRabies2019; @sreenivasanOverviewRabiesPostexposure2019; @tarantolaRabiesPostexposureProphylaxis2017]. Currently, each clinic in Madagascar serves on average a catchment population of 750,000 persons {inline}. And while PEP is free, KAP bit?  In fact, 22 clinics still charge patients (on average ~ 2000 Ar but anywhere from 500 - 10000 Ar for costs other than the vaccine) so this could be a factor too. Latin American countries, where significant progress has been made towards elimination, aim to provide one clinic providing PEP for 100,000 persons, which would require approximately 200 additional ARMC in Madagascar. If GAVI investment results in expansion to underserved areas and strengthens existing provisioning and supply chain mechanisms, then rabies burden could be greatly reduced. Importantly, our estimates of expansion suggest that it remians cost-effectivehowever these remain reasonably cost-effective. Just given vial costs alone, the cost would be between xxx and yyy per patient (similar to GAVI). Given a GAVI investment, if the government could shift to improving provisioning systems. Cost effectiveness point here (not per DALY but per death averted.)

Ultimately even given expanded access, achieving 'the last mile', i.e. preventing deaths due to rabies in the most remote populations will require a disproportionate level of resources and our results suggest that this will not be achievable through PEP alone [@delriovilasTribulationsLastMile2017]. IBCM programs and risked based assessments are one way to tackle the issue of managing PEP resources effectively even as demand outpaces reductions in burden [@rysavaPathRabiesElimination2018]. Combining PEP provisioning with mass dog vaccination will be key to interrrupting transmission, and eliminating deaths due to rabies. These same issues of access, however apply to both dog vaccination programs and surveillance systems, and integrating spatial heterogeneities in vaccination and detection will be critical to understand how control and prevention interventions can be best implemented [@fergusonHeterogeneitySpreadControl2015]. Wentworth human right bit...

## Conclusion {inline these stats}
Adding clinics alone wont be enough. Need to improve awareness & clinic functioning. 
Conclusions = You should expand access. Focusing on expanding access to a set of clinics that can reliable provision PEP might be the best bet (rather than rolling out to all clinics). Developing novel methods of distributing PEP/decentralized way, i.e. drone delivery. Data is really important at all levels! Clinic functioning, health seeking behavior, and also acting as sentinels for rabies surveillance. But you won't get rid of deaths completely this way.

Geographic access to PEP is a strong predictor of reported bite incidence in Madagascar and expanding PEP availability to more clinics across the country could greatly reduce human rabies deaths. Based on our modeling framework, a more equitable target may be aiming to minimize the population living within 3-5 hours travel time of a clinic, but due to broad patterns of remoteness across Madagascar, even when all primary health facilities are providing PEP, approximately 10 % {inline} of people live > 3 hours away from a clinic. Expanding PEP to clinics that target currently underserved populations may be the best strategy for preventing human rabies deaths. If expanded only to one clinic per district (N = 114), an additional 40% of deaths could be averted. Expansion of PEP can greatly reduce risk and inequities in access to this life-saving vaccine, but ultimately, interrupting transmission of rabies through mass dog vaccination programs will be the most equitable way to prevent, and eventually eliminate, rabies deaths.

## Acknowledgements

\newpage

## References
::: {#refs_main}
:::

\newpage

# SUPPLEMENTARY FIGURES & TABLES

## S1. Estimating travel times to the closest ARMC {#S1}
[*jump back to main results section: estimating travel times*](#results-ttimes)

```{r fig S1.1, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S1.1_inputs.jpeg"))
```
*__[Figure S1.1.]{#figS1.1}__ Raster inputs to estimate travel times to the closest ARMC for Madagascar __A)__ Friction surface of travel speeds (in minutes per meter) at an ~ 1x1 km scale __B)__ Population estimates resampled to the same friction surface.*

```{r fig S1.2, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S1.2_raw.jpeg"))
```
*__[Figure S1.2.]{#figS1.2}__ Source of true travel time estimates __A)__ Distribution of travel times estimated at the grid cell level and reported by patients for each commune where patient data were available from the Moramanga ARMC. __B)__ Reported driving times between locations, with the color corresponding to the total driving time and the size of the line showing the direction of travel (narrow to wide ~ origin to destination). Paths are bezier curves from origin to destination, but do not show actual paths driven.*

```{r fig S1.3, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S1.3_modes.jpeg"))
```
*__[Figure S1.3.]{#figS1.3}__ Reported modes of transport used compared to reported travel times for patients reporting to the Moramanga ARMC.*

```{r fig S1.4, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S1.4_lms.jpeg"))
```
*__[Figure S1.4.]{#figS1.4}__ Observed estimates of travel times (commune means of patient reported travel times and driving times between point locations) predicted by __A)__ Distance (km) (Euclidean distance between origin and destination for driving times and distance from the commune centroid to the Moramanga ARMC for commune means) __B)__ Travel time estimates and C) Travel time estimates weighted by population (for commune means only).*

*__[Table S1.1.]{#tabS1.1} R^2 values from linear models with estimated access metrics predicting either driving times or commune means of patient reported travel times.*
```{r table S1.1}
S1.1_table <- read.csv(here("output/stats/access_met_r2.csv"), row.names = 1)
kable(S1.1_table, col.names = c("Predictor", "Driving times", 
                                "Commune mean of patient reported travel times"))
```

```{r fig S1.5, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S1.5_outputs.jpeg"))
```
*__[Figure S1.5.]{#figS1.5}__ Estimates of mean travel times weighted by population at the __A)__ District __B)__ Commune scale.*

## S2. Estimating average annual bite incidence at the district and commune level
[*jump back to main results section: Bite patient data*](#results-biteinc)

```{r fig S2.1, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S2.1_catchments.jpeg"))
```
*__[Figure S2.1]{#figS2.1} __A)__ Catchments as assigned by closest clinic for the majority of the population within a commune (polygon fill) or which a district (polygon outline). Admin units where the fill and border colors do not match show places where assigned catchments differ at the district vs commune scale. __B)__ Distribution of the proportion of the population in a given administrative unit (district or commune) served by the catchment assigned. __C)__ The proportion of bites reported to each clinic which originated from a district within the assigned catchment.*

```{r fig 2.2, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S2.2_vialcomps.jpeg"))
```
*__[Figure S2.2.]{#figS2.2}A)__ The daily time series of the number of forms submitted by each clinic, with periods of time where no forms were submitted for >= 15 days excluded (in white). __B)__ Estimates for the proportion of forms submitted for each clinic (points are the estimate for each year) calculated as the # of days in a year which were not excluded based on the  criteria of 15 consecutive days of non-submission/365. __C)__ The difference between log(estimated) and log(observed) vials provisioned for the period of 2013-2016 for each clinic correcting for undersubmission (squares) using the 15 day cut-off show in A and B, vs. not correcting for undersubmission (circles). We did not have data on vials provisioned for IPM.*

## S3. Modeling reported bite incidence as a function of travel times {#S3}

[*jump back to main results section: Travel times as a predictor of bites*](#results-bitemods)

```{r fig S3.1, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S3.1_corrs.jpeg"))
```
_**Figure S3.1.** Correlation between travel time in hours and population size of administrative unit the district and commune scale._

<br/>

*__Table S3.1.__ DIC Estimates for all models. Models with the lowest DIC for each dataset and scale of covariate are in bold.  For the column pop effect, addPop = models with population size as additional covariate, onlyPop = models with population as only covariate, flatPop = models with population as offset in model. For the intercept type: random = random intercept by catchment, fixed = a single fixed intercept was estimated).*
```{r table 1}
model_ests <- read.csv(here("output/mods/estimates.csv"))
model_ests %>%
  select(params, Mean, pop_predict, intercept, data_source, scale) %>%
  spread(key = params, value = Mean, fill = 0) -> model_means
model_ests %>%
  group_by(data_source, scale, pop_predict, intercept) %>%
  summarize(dic = mean(dic)) -> dic_ests
dic_ests %>%
  group_by(data_source) %>%
  arrange(dic, .by_group = TRUE) %>%
  mutate(dic = round(dic, 2)) -> dic_ranks
knitr::kable(dic_ranks, 
             col.names = c("Dataset", "Scale", "Pop effect", "Intercept type", "DIC"))
```
<br/>

```{r fig S3.2, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S3.2_fitted.jpeg"))
```
_**Figure S3.2. Prediction to data used to fit each model.** Log of the observed bites against the log of predicted bites using the mean of the posterior distributions for each parameter. Columns are by the type of model intercept (either a fixed intercept or a random intercept by catchment) and rows are the type of model structure with respect to the population covariate (addPop = population size as additional covariate, onlyPop = population as only covariate, flatPop = population as offset in model). Colors show which data set  was used for fitting and the scale of the model (Moramanga = Moramanga data with covariates at the commune level, Commune = National data with covariates at the commune level, District = National data with covariates at the district level)._\

<br/>

```{r fig S3.3, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S3.3_outfit.jpeg"))
```
_**Figure S3.3. Out of fit predictions to data.** Log of the observed bites against the log of predicted bites generated from the mean of the posterior distributions from each model for data not used to fit the model. The first two columns are the predictions from the commune and district model fitted to the national data for the Moramanga data with fixed and random intercepts. The third column are predictions from models fitted to the Moramanga data for the national data at the commune and district scale (only fixed intercept models). Rows are the type of model structure with respect to the population covariate and colors show which data set was used for fitting as per Figure S3.2._

<br/>

```{r fig S3.4, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S3.4_ODcheck.jpeg"))
```
_**Figure S3.4.** Overdispersion....

```{r fig S3.5, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S3.5_cutoffs.jpeg"))
```
_**Figure S3.5.** The estimated relationship between travel time in hours (x-axis) and mean annual bite incidence per 100,000 persons (y-axis) for models with a random catchment intercept and with population as an offset. Panels show predictions from models fitted to (1) data corrected for both undersubmission by correcting for periods of at least 7 days where zero patient forms where submiited, (2) correcting for periods of 15 days where zero patient forms where submiited, and (3) with the raw data not correcting for undersubmission. The lines are the estimates using the mean of the posterior distributions of the parameters and the envelopes shows the 95% CI given the random intercept. The colors are the two candidate models with travel times at the commune and district level fitted to the national data with a random intercept by catchment. The points show the data used to fit the models._

## S4. Range of rabies exposure incidence in people

```{r fig S4.1, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S4.1_exprange.jpeg"))
```

_**Figure S4.1.** Estimated exposures per 100,000 persons using parameters ranges from the literature and data from the Moramanga District. The vertical bars show the range of human to dog ratios (HDRs) observed in the literature across African countries. The dashed lines show the HDRs estimated from the Moramanga district from a recent study. The cells with red outlines show the range of estimated exposure incidence from a previous study of bite patients in Mormanga District. For the range of exposure incidence used to estimate burden in this study, we used a minimum consistent with higher hdr and a maximum consistent with an upper hdr and set the midpoint to be the lower end of our estimates from Moramanga so that it makes sense with previous estimates & assumptions. However in the se analysis we looked at the upper range of this as well....._

## S5. Estimating the impact of expanding PEP provisioning to additional clinics {#S5}
[*jump back to main methods section: Incremental reduction of burden*](#methods-incremental)  
[*jump back to main results section: Incremental reduction of burden*](#results-incremental)

<!-- TO DO: add catchmaps -->

```{r fig S5.1, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.1_ARMCadded.jpeg"))
```
_**Figure S5.1.** Map of the location and at what step each clinic was added. The circles are each of the CSB II across the country sized by the proportion of the population for which they reduced travel times (excluding those already living within 3 hours of an ARMC) and the colors showing at what step each clinic was added (i.e. 1 = the first clinic added). The large white crosses show the location of the existing 31 ARMC in Madagascar and the smaller white crosses are the remaining CSB II in the country (which reduced travel times for less than 0.01% of people living greater than 3 hours from a clinic and thus were not ranked)._

```{r fig S5.2, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.2_map_shifts.tiff"))
```
_**Figure S5.2.** Maps of how travel times change as clinics are added at the **A)** ~1 x 1 km grid cell **B)** commune and **C)** district scales. The columns are ordered by the number or clinics at each step: baseline (N = 31), 100, 200, 300, 723 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country). Grey pixels show the location of ARMC at each step. Commune and district values are the average grid cell travel times weighted by the population in each cell._

```{r fig S5.3, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.3_prop_ttimes.jpeg"))
```
_**Figure S5.4.** As clinics are added, the proportion of the population binned by travel times in hours. Panel labels are the number or clinics at each step: baseline (N = 31), 100, 200, 300, 472 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country)._

```{r fig S5.4, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.4_prop_catch.jpeg"))
```
_**Figure S5.3.** As clinics are added, shifts in distribution of the proportion of population in a given catchment  for the two models (colors, Commune scale model fitted to the Moramanga data and District scale model fitted to the National data). Rows are ordered by the number or clinics at each step: baseline (N = 31), 100, 200, 300, 472 (when algorithm stops adding clinics), and max (N = 1648 clinics, all CSB IIs in the country)._

```{r fig S5.5, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.5_ttimes_shifts.jpeg"))
```
_**Figure S5.5.** Shifts in estimated **A)** travel times (hrs), **B)** bite incidence per 100,000 persons. The left panels show the shift in distributions of these estimates as clinics are added, and the right panels show the changes in the averages (line) and the 95% quantiles (shading) for communes and districts  for the two models (colors, Commune scale model fitted to the Moramanga data and District scale model fitted to the National data)._

```{r fig S5.6, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.6 _reporting_shifts.jpeg"))
```
_**Figure S5.6.** Shifts in estimated **A)** Reporting, **B)** death incidence per 100,000 persons. The left panels show the shift in distributions of these estimates as clinics are added, and the right panels show the changes in the averages (line) and the 95% quantiles (shading) for communes and districts for the two models (colors, Commune scale model fitted to the Moramanga data and District scale model fitted to the National data)._

```{r fig S5.7, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.7_clinic_shifts.jpeg"))
```
_**Figure S5.7.** Shifts in clinic average **A)** catchment population size, **B)** and annual vial demand, and **C)**  daily throughput given estimates of bite incidence for the two models (colors, Commune scale model fitted to the Moramanga data and District scale model fitted to the National data). The left panels show the shift in distributions of these estimates as clinics are added, and the right panels show the changes in the averages (line) and the 95% quantiles (shading). Catchment population sizes are the same for each model._

```{r fig S5.8, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.8_comm_map.jpeg"))
```
_**Figure S5.8.** Where bites are reported to as clinics are added for the Commune scale model fitted to the Moramanga data. The color of the points show at what step the clinic was added (with the grey points in the last panel the CSB II in the country that reduced travel times for less than 0.01% of people living greater than 3 hours from a clinic and thus were not ranked), the size shows the average number of bites reported to that clinic annually, and the lines show from which commune bites are reported from. The shading of the polygons show the reported bite incidence at the commune level._

```{r fig S5.9, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.9_dist_map.jpeg"))
```
_**Figure S5.9.** Where bites are reported to as clinics are added for the District scale model fitted to the National data. The color of the points show at what step the clinic was added (with the grey points in the last panel the CSB II in the country that reduced travel times for less than 0.01% of people living greater than 3 hours from a clinic and thus were not ranked) and the lines show from which district bites are reported from. The shading of the polygons show the reported bite incidence at the district level._

```{r fig S5.10, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S5.10_prop_in_admin.jpeg"))
```
_**Figure S5.9.** Shifts in the distribution of the proportion of bites reported to a clinic that came from within the same A) commune or B) district as the ARMC was located for the two models (Commune scale model fitted to the Moramanga data and District scale model fitted to the National data)._

## S6. Sensitivity of burden estimates to parameter assumptions {#S6}

[*jump back to main methods section: Incremental reduction of burden*](#methods-sensitivity)  
[*or to main results section: Incremental reduction of burden*](#results-sensitivity)

```{r fig S6.1, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S6.1_base_se.jpeg"))
```
_**Figure S6.1.** ._

```{r fig S6.2, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S6.2_addARMC_se.jpeg"))
```
*__Figure S6.2.__ Predicted relationship between travel times and rabies death incidence per 100k persons for the district model vs. commune models (colors) with columns arranged by the scaling factor (minimum incidence of 11.1 rabies exposures/100,000 persons, the maximum negative scaling with population size, the maximum positive scaling with population size, and maximum incidence of 55.7 rabies exposures/100,000 persons, see Fig S6.1 for the scaling factors). Rows show burden estimates given the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]). Colors represent the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]); shapes show the maximum and minimum assumption of $\rho_{max}$ = [0.8 , 0.95]).*

```{r fig S6.3, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S6.3_vial_se.jpeg"))
```
*__Figure S6.3.__ Estimates of the impact of expanding access to PEP incrementally on  burden at the national level for the district model vs. commune models (colors) with columns arranged by the scaling factor (minimum incidence of 11.1 rabies exposures/100,000 persons, the maximum negative scaling with population size, the maximum positive scaling with population size, and maximum incidence of 55.7 rabies exposures/100,000 persons--see Fig S6.1 for the scaling factors). Rows show burden estimates given the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]). Colors represent the maximum and minimum assumption of the proportion of reported bites that are assumed to be rabid ($p_{rabid}$ = [0.2, 0.6]); shapes show the maximum and minimum assumption of $\rho_{max}$ = [0.8 , 0.95].*

```{r fig S6.4, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S6.4_scaling_rels.jpeg"))
```
*__Figure S6.4.__ Range of constrained scaling factors generated for district and commune population size. The black dashed lines represent the minimum and maximum estimated rabies exposure incidence (minimum = 11.1 exposures/100k persons, maximum = 55.7 exposures/100k persons). We generate constrained scaling factors with human population size so that rabies exposure incidence in each location scales positively or negatively across the range of observed population sizes (panels)*

```{r fig S6.5, fig.fullwidth=TRUE}
include_graphics(here("figs/supplementary/S6.5_scaling_se.jpeg"))
```
*__Figure S6.5.__ *

## Supplementary Reference (R packages used) {#srefs}

::: {#refs_R}
:::