### 2021-03-08 analysis/paper/subfiles/cover_letter.Rmd

---
title: ""
author: ""
date: "20/02/2020"
output: 
  pdf_document:
    keep_tex: false
---

Dear Editor,

We are submitting our manuscript entitled “Lessons learned and paths forward for rabies dog vaccination in Madagascar: a case study of pilot vaccination campaigns in Moramanga District” for consideration for publication in Tropical Medicine and Infectious Disease (particularly for consideration for the Special Issue on One Health and NTDs, per communication between editors and co-author K. Hampson). Our article presents results from two pilot dog rabies vaccination campaigns in the Moramanga District and integrates the data we collected during the campaign into quantitative analyses of costs, coverage, and potential future vaccination strategies. We feel that this work will be very useful for not only Madagascar, but sub-Saharan African countries in general hoping to scale up dog vaccination, and to one health practioners more broadly in thinking about how best to translate strategies for public health implementation across human and animal sectors.

We thank you for your consideration.

Sincerely,  
Malavika Rajeev, Princeton University